---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

tidy-select

# - *fundamentals*

When we apply the same function to different columns,

```{r}
df %>%
  summarise(Avg_Quantity = mean(Quantity),
            Avg_Price = mean(Price))
```

we can use `across()` to produce more concise and less prone to typos code.

```{r}
df %>% 
  summarise(across(c(Quantity, Price), mean))
```

In its first argument we specify the columns we want to apply the function to, using the same syntax as `select()`.

```{r, warning = FALSE}
df %>% 
  summarise(across(c(Quantity, Price), mean))
df %>% 
  summarise(across(c(4, 6), mean))
df %>% 
  summarise(across(Quantity:Price, mean))
df %>% 
  summarise(across(where(is.numeric), mean))
df %>%
  summarise(across(c(where(is.numeric), -`Customer ID`), mean))
df %>%
  summarise(across(starts_with(c("Q", "P")), mean))
df %>%
  summarise(across(everything(), mean))
```

And in the second the function we want to apply, that can be written in other ways as well,

```{r}
df %>%
  summarise(across(where(is.numeric), ~ mean(.x)))
df %>%
  summarise(across(where(is.numeric), function(x) mean(x)))
```

which is useful in case the function has optional arguments we want to employ.

```{r}
df %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))
df %>%
  summarise(across(where(is.numeric), function(x) mean(x, na.rm = TRUE)))
```

`across()` can be used several times in the same call.

```{r}
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(across(where(is.numeric), mean),
            across(where(is.character), n_distinct),
            n = n())
```

# - *.names*

With the `.names` argument we can change the output names from the default `{.col}` (which is the name of the input columns).

```{r}
df %>%
  summarise(across(c(Quantity, Price), mean))
df %>%
  summarise(across(c(Quantity, Price), mean, .names = "Avg_{.col}"))
```

It uses the `glue` library syntax and it allows for important modifications.

```{r}
df %>%
  summarise(across(c(Quantity, Price), mean, .names = "avg_{stringr::str_to_lower(.col)}"))
```

If we are using `mutate()` though we need to be careful as the lack of the `.names` argument overwrites the original columns.

```{r}
df %>%
  mutate(across(c(Quantity, Price), mean), .keep = "used")
df %>%
  mutate(across(c(Quantity, Price), mean, .names = "Avg_{.col}"), .keep = "used")
```

`across()` creates tibbles, so using it in the following way will add a column with a tibble for every row.

```{r}
df %>%
  mutate(Total_Expense = across(c(where(is.numeric), -`Customer ID`), mean), .keep = "used")
```

The results inside every tibble are the same.

```{r}
df %>%
  mutate(Total_Expense = across(c(where(is.numeric), -`Customer ID`), mean), .keep = "used") %>%
  select(Total_Expense) %>%
  slice(1:2) %>%
  tidyr::unnest(cols = c(Total_Expense))
```

To not get confused by all the parenthesis, we stress the fact that `across()` is the sole argument of the `summarise()`/`mutate()`/etc. call.

```{r}
df %>%
  summarise(
    across(c(Quantity, Price), ~ mean(.x, na.rm = TRUE), .names = "avg_{stringr::str_to_lower(.col)}")
    )
```

# - *multiple functions in an external list*

Instead of just one function, we can use several if we put them in a list.

```{r, df_print = "paged"}
(avg_std <- list(avg = ~ mean(.x),
                 st_dev = ~ sd(.x)))
```
```{r}
df %>%
  summarise(across(c(Quantity, Price), avg_std))
```

Take notice of the names of the output columns that stem from the default `.names = {col}_{fn}` for when we use a list of functions.
As before, that can be modified with a `glue` syntax and the appropriate strings manipulation functions.

```{r}
df %>%
  summarise(across(c(Quantity, Price), avg_std, .names = "{col}.{stringr::str_to_title(fn)}"))
df %>%
  summarise(across(c(Quantity, Price), avg_std, .names = "{stringr::str_to_title(fn)}.{col}"))
```

In case we want to have akin columns close to each other, we have to split the code in two (or piping another `select()`/`relocate()` call).

```{r}
df %>%
  summarise(across(c(Quantity, Price), ~ mean(.x), .names = "Avg.{col}"),
            across(c(Quantity, Price), ~ sd(.x), .names = "St_dev.{col}"))
```

When the list is not named,

```{r, df_print = "paged"}
(avg_std_2 <- list(~ mean(.x),
                   ~ sd(.x)))
```

the default is to differentiate the columns by numbering them.

```{r}
df %>%
  summarise(across(c(Quantity, Price), avg_std_2))
```

Using a list of functions with `mutate()` adds new columns at the end of the data frame. 

```{r}
df %>%
  mutate(across(c(Quantity, Price), avg_std))
df %>%
  mutate(across(c(Quantity, Price), avg_std_2))
```

When we use a function that returns more than one output it's better to use `reframe()` instead of `summarise()`.

```{r}
df %>%
  reframe(across(c(Quantity, Price), range))
```

And when it is used together with single output functions the result of the latter will be replicated.

```{r, df_print = "paged"}
(range_iqr <- list(range = ~ range(.x),
                   iqr = ~ IQR(.x)))
```
```{r}
df %>%
  reframe(across(c(Quantity, Price), range_iqr))
```

# - *possible selection issues*

`across()` can also evaluates columns just created if they answer to the selection statement, like the following numeric `n`, giving undesired results, like its standard deviation (which is a single number, hence the NA) instead of the number of rows of the data frame.

```{r}
df %>%
  summarise(n = n(), 
            across(c(where(is.numeric), -`Customer ID`), sd))
```

There are several solutions to that:  
we can move `n` after the `across()` call, in this way it will not be evaluated by it

```{r}
df %>%
  summarise(across(c(where(is.numeric), -`Customer ID`), sd), 
            n = n())
```

we can explicitly exclude `n` from the selection if we care about the order of the output

```{r}
df %>%
  summarise(n = n(), 
            across(c(where(is.numeric), -c(`Customer ID`, n)), sd))
```

or we can wrap everything in a data frame where `n` is independent from the columns returned by `across()`.

```{r}
df %>% 
  summarise(tibble(n = n(), 
                   across(c(where(is.numeric), -`Customer ID`), sd)))
```

# - *cur_column()*

`cur_column()` is an `across()` specific function that returns the column that is currently being evaluated.

```{r}
df %>% 
  mutate(across(c(Quantity, Price), ~ paste(cur_column(), .x)), .keep = "used")
```

That line of code is equivalent to this one, where we can see that it returns the name of the column, not the values.

```{r}
df %>%
  mutate(Quantity = paste("Quantity", Quantity),
         Price = paste("Price", Price), .keep = "used")
```

We can taking advantage of `cur_column()` if for example we want to multiply several numeric columns for a different number each.

To do so we have to beforehand define a named list with the numbers we want to multiply every column with,

```{r, df_print = "paged"}
(mult_list <- list(Quantity = 2,
                   Price = 10))
```

and then in the function argument of `across()` we can access that specific number by subsetting the list with `cur_column()`.

```{r}
df %>% 
  mutate(across(c(Quantity, Price), ~ .x * mult_list[[cur_column()]], .names = "Mult_{.col}"), .keep = "used")
```

We can also use it to dynamically change the arguments of the function.

```{r, df_print = "paged"}
(n_digital <- list(Quantity = 2,
                   Price = 1))
```
```{r}
df %>% 
  mutate(across(c(Quantity, Price), ~ round(.x, n_digital[[cur_column()]]), .names = "Rounded_{.col}"), .keep = "used")
```

If the columns selected are many we can access them more concisely using the selection helper `all_of()`.

```{r}
df %>%
  mutate(across(all_of(names(mult_list)), ~ .x * mult_list[[cur_column()]], .names = "Mult_{.col}"), .keep = "used")
df %>%
  mutate(across(all_of(names(n_digital)), ~ round(.x, n_digital[[cur_column()]]), .names = "Rounded_{.col}"), .keep = "used")
```

# - *.unpack*

In case the function returns data frames,

```{r}
df %>%
  mutate(across(c(Quantity, Price), ~ tibble(Min = min(.x), Max = max(.x)), .names = "MinMax_{.col}"), .keep = "used")
```

we can use `.unpack = TRUE` to expand them.

```{r}
df %>%
  mutate(across(c(Quantity, Price), ~ tibble(Min = min(.x), Max = max(.x)), .unpack = TRUE), .keep = "used")
```

And, as with `.names`, we can modify the output from the default `{outer}_{inner}`, where the former refer to the names of `df` and the latter to the names of the data frames created by the function.

```{r}
df %>%
  mutate(across(c(Quantity, Price), ~ tibble(Min = min(.x), Max = max(.x)), .unpack = "{inner}_{outer}"), .keep = "used")
```

# - *not applicable verbs*

`across()` doesn't work with other columns manipulation verbs that use the same tidy-select syntax, like `select()`, `rename()` and `relocate()`.

```{r, error = TRUE}
df %>%
  select(across(c(Quantity, Price)))
```

Instead of `rename()` we can use `rename_with()` in case we want to change the names of columns.
Notice how the two arguments (the selection of columns and the function to be applied to them) are in a different order compared to `across()`.

```{r}
df %>%
  rename_with(stringr::str_to_lower, c(Quantity, Price))
```

# - *other applicable verbs*

`across()` works though with other rows manipulation verbs that uses the `mutate()` data-masking syntax.
Except for `slice()`, as that verb doesn't have a selection of several columns between its arguments.

## - *arrange()*

With `arrange()` it can speed up typing if we are using columns with similarities to take advantage of.

```{r}
df %>%
  arrange(across(starts_with("I"), desc))
df %>%
  arrange(desc(Invoice), desc(InvoiceDate))
```

We have no way of individually control the order of the columns though.

```{r}
df %>%
  arrange(desc(Invoice), InvoiceDate)
```

## - *filter() (with if_all()/if_any())*

With `filter()` we can use `across()` to facilitate the following code, where we preserve the rows that have a value greater than 5 in all the columns specified.

```{r}
df %>%
  filter(Quantity > 5 &
           Price > 5)
df %>%
  filter(across(c(Quantity, Price), ~ .x > 5))
```

But, as the warning suggests, it is better to use `if_all()` as `across()` for `filter()` is deprecated.

```{r}
df %>%
  filter(if_all(c(Quantity, Price), ~ .x > 5))
```

In case we want to preserve rows that have a value greater than 5 in any of the columns selected,

```{r}
df %>%
  filter(Quantity > 5 |
           Price > 5)
```

we are obliged to use `if_any()`, as `across()` is not able to do that.

```{r}
df %>%
  filter(if_any(c(Quantity, Price), ~ .x > 5))
```

We can also use them to filter for the same interval on several columns, either on all of them at the same time,

```{r}
df %>%
  filter(between(Quantity, 10, 20) &
           between(Price, 10, 20))
df %>%
  filter(if_all(c(Quantity, Price), ~ between(.x, 10, 20)))
df %>%
  filter(if_all(c(Quantity, Price), ~ .x >= 10 & .x <= 20))
```

or just on any of them.

```{r}
df %>%
  filter(between(Quantity, 10, 20) |
           between(Price, 10, 20))
df %>%
  filter(if_any(c(Quantity, Price), ~ between(.x, 10, 20)))
df %>%
  filter(if_any(c(Quantity, Price), ~ .x >= 10 & .x <= 20))
```

By design though they don't understand `cur_column()` so if we want to filter each column for a different value 

```{r}
df %>%
  filter(Quantity > 2 &
           Price > 10)
```

we need to use `across()` (thus an `&` statement), therefore we can only preserve the rows where both conditions are TRUE at the same time.

```{r, df_print = "paged"}
(fltr <- list(Quantity = 2,
              Price = 10))
```
```{r}
df %>%
  filter(across(c(Quantity, Price), ~ .x > fltr[[cur_column()]]))
```

Another example with `between()`, where we filter on two different ranges.

```{r}
df %>%
  filter(between(Quantity, 10, 20) &
           between(Price, 20, 30))
```
```{r, df_print = "paged"}
(fltr2 <- list(Quantity = c(10, 20),
               Price = c(20, 30)))
```
```{r}
df %>%
  filter(across(c(Quantity, Price), ~ between(.x, fltr2[[cur_column()]][1], fltr2[[cur_column()]][2])))
```

`if_all()` and `if_any()` can also be used in `mutate()` or `summarise()` calls, as long as we exploit their property of returning logical vectors.

```{r}
df %>%
  mutate(High = case_when(Quantity > 50 & Price > 50  ~ "volume_and_price",
                          Quantity > 50 | Price > 50  ~ "volume_or_price",
                          Quantity < 50 & Price < 50  ~ "neither"), .keep = "used")
df %>%
  mutate(High = case_when(if_all(c(Quantity, Price), ~ .x > 50) ~ "volume_and_price",
                          if_any(c(Quantity, Price), ~ .x > 50) ~ "volume_or_price",
                          if_all(c(Quantity, Price), ~ .x <= 50) ~ "neither"), .keep = "used")
```

## - *group_by()*

`across()` can be used with `group_by()` as well, with just the first argument.

```{r}
df %>%
  group_by(across(c(Quantity, Price)))
df %>%
  group_by(across(everything())) %>%
  filter(n() > 1)
```

## - *count(), distinct() & other verbs*

That worked because `across()` returns a data frame, so we can utilize `across()` with every function that takes one, like `count()`,

```{r}
df %>%
  count(across(c(Country, `Customer ID`)))
```

or `distinct()`.

```{r}
df %>%
  distinct(across(c(Country, `Customer ID`)))
```

But that doesn't mean that we can't apply a function as well when it is appropriate.

```{r}
df %>%
  distinct(across(c(Country, `Customer ID`),  ~ stringr::str_to_lower(.x)))
```

If the data frame already has logical columns,

```{r}
df %>%
  mutate(Boolean_TFT = rep(c(TRUE, FALSE, TRUE), length.out = nrow(df)),
         Boolean_TF = rep(c(TRUE, FALSE), length.out = nrow(df)), .before = "Invoice")
```

`if_all()` and `if_any()` too can be used without a function.

```{r}
df %>%
  mutate(Boolean_TFT = rep(c(TRUE, FALSE, TRUE), length.out = nrow(df)),
         Boolean_TF = rep(c(TRUE, FALSE), length.out = nrow(df)), .before = "Invoice") %>%
  filter(if_all(where(is.logical)))
df %>%
  mutate(Boolean_TFT = rep(c(TRUE, FALSE, TRUE), length.out = nrow(df)),
         Boolean_TF = rep(c(TRUE, FALSE), length.out = nrow(df)), .before = "Invoice") %>%
  filter(if_any(where(is.logical)))
```

# - *with group_by()*

With grouped data frames the grouping columns are never included in the selection, even if they satisfy the selecting condition.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  summarise(across(where(is.numeric), mean))
```

## - *cur_group()*

We've seen previously that `cur_column()` accesses the currently selected column.   
`cur_group()` lets us access the current group, so it can be used if we want to modify each row for a value specific to the group it pertains.

```{r, df_print = "paged"}
(mod_tbl <- tibble("France" = 2,
                   "Italy" = 10,
                   "Germany" = 5))
```
```{r}
df %>%
  filter(Country %in% c("France", "Italy", "Germany")) %>%
  select(Country, Quantity, Price)
df %>%
  filter(Country %in% c("France", "Italy", "Germany")) %>%
  select(Country, Quantity, Price) %>%
  group_by(Country) %>%
  mutate(across(c(Quantity, Price), ~ .x * mod_tbl[[cur_group()$Country]], .names = "Mod_{.col}"))
```

Compared to before, notice how we needed to define a tibble, and not a list, for the `Country` specific values and also to subset `cur_group()` with the name of the grouping column.

Using `cur_column()` and `cur_group()` jointly, we can also modify each column for a different value depending on the group.

```{r, df_print = "paged"}
(mod_tbl2 <- tibble("Column" = c("Quantity", "Price"),
                    "France" = c(2, 10),
                    "Italy" = c(10, 5),
                    "Germany" = c(5, 2)))
```
```{r}
df %>%
  filter(Country %in% c("France", "Italy", "Germany")) %>%
  select(Country, Quantity, Price) %>%
  group_by(Country) %>%
  mutate(across(c(Quantity, Price), ~ .x * unlist(mod_tbl2[mod_tbl2$Column == cur_column(), cur_group()$Country]), .names = "Mod_{.col}"))
```

It is important that we reduce the tibble to a vector with `unlist()`, otherwise `across()` would return list columns of data frames,

```{r}
df %>%
  filter(Country %in% c("France", "Italy", "Germany")) %>%
  select(Country, Quantity, Price) %>%
  group_by(Country) %>%
  mutate(across(c(Quantity, Price), ~ .x * mod_tbl2[mod_tbl2$Column == cur_column(), cur_group()$Country], .names = "Mod_{.col}"))
```

that we could unpack, but they don't present the desired output (for every group, the first row is recycled along its size).

```{r}
df %>%
  filter(Country %in% c("France", "Italy", "Germany")) %>%
  select(Country, Quantity, Price) %>%
  group_by(Country) %>%
  mutate(across(c(Quantity, Price), ~ .x * mod_tbl2[mod_tbl2$Column == cur_column(), cur_group()$Country], .names = "Mod_{.col}", .unpack = TRUE))
```

# - *using pick() instead*

tidy-select

As we've seen before, `across()` can be used with the function argument left empty.   
In this way it will just return, with specific verbs, the columns selected.

```{r}
df %>% 
  reframe(across(c(Quantity, Price)))
df %>% 
  mutate(across(c(Quantity, Price)))
```

For these tasks is better to use `pick()` though, because, being that its only purpose, we can list columns without wrapping them in `c()`.

```{r}
df %>% 
  reframe(pick(Quantity, Price))
df %>% 
  mutate(pick(Quantity, Price))
```

Like `across()`, `pick()` returns a data frame, so it can be used to feed one into a function.

```{r}
df %>%
  group_by(pick(starts_with("c")))
df %>%
  distinct(pick(2, 3))
df %>%
  count(pick(1, StockCode))
df %>%
  slice_min(pick(Quantity, StockCode), n = 4)
```

And likewise with grouped data frames the grouping column is never included in the selection, even if it satisfies the selecting condition.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  count(pick(where(is.numeric)))
df %>%
  count(pick(where(is.numeric)))
```

`pick()` doesn't have any optional arguments.