---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

data-masking

# - *case_when()*

We can nest several if_else()s if we want a variable with more than two classes in the output.

```{r, df_print = "paged"}
EU <-  c("Austria", "Belgium", "Channel Islands", "Cyprus", "Denmark", "EIRE", "Finland", "France", "Germany", 
         "Greece", "Iceland", "Italy", "Lithuania", "Malta", "Netherlands", "Norway", "Poland", 
         "Portugal", "Spain", "Sweden", "Switzerland")
```
```{r}
df %>%
  mutate(in_EU = if_else(Country == "Unspecified", Country, 
                         if_else(Country %in% EU, "Yes", "No")), .keep = "used")
```

But it is best to use case_when(), which is built for this purpose.
It uses a formula like syntax, where on the right of ~ we specify which value to assign when the test on the left evaluates to TRUE.

```{r}
df %>%
  mutate(in_EU = case_when(Country == "Unspecified" ~ Country,
                           Country %in% EU ~ "Yes",
                           !Country %in% EU ~ "No"), .keep = "used")
```

We can't specify any value for when a test evaluates to FALSE so, when an element doesn't evaluate to TRUE to any test, the output is NA.

```{r}
df %>%
  mutate(in_EU = case_when(Country == "Unspecified" ~ Country,
                           Country %in% EU ~ "Yes"), .keep = "used")
```

This means that each element needs a test that evaluates to TRUE, in order to not produce NAs. 

## - *.default*

Sometimes that can be bothersome to do, like if we only wanted to modify the name of 2 elements, and we might also miss on some values.

```{r}
df %>%
  mutate(Country = case_when(Country == "United Kingdom" ~ "UK",
                             Country == "EIRE" ~ "Ireland",
                             Country == "Austria" ~ Country,
                             Country == "Belgium" ~ Country,
                             Country == "Channel Islands" ~ Country,
                             Country == "Cyprus" ~ Country,
                             Country == "Denmark" ~ Country,
                             Country == "Finland" ~ Country,
                             Country == "France" ~ Country,
                             Country == "Germany" ~ Country,
                             Country == "Greece" ~ Country,
                             Country == "Iceland" ~ Country,
                             Country == "Italy" ~ Country,
                             Country == "Lithuania" ~ Country,
                             Country == "Malta" ~ Country,
                             Country == "Netherlands" ~ Country,
                             Country == "Norway" ~ Country,
                             Country == "Poland" ~ Country,
                             Country == "Portugal" ~ Country,
                             Country == "Spain" ~ Country,
                             Country == "Sweden" ~ Country,
                             Country == "Switzerland" ~ Country), .keep = "used")
```

But we can use the .default argument for assigning a value of choice instead of NA.

```{r}
df %>%
  mutate(Country = case_when(Country == "United Kingdom" ~ "UK",
                             Country == "EIRE" ~ "Ireland", .default = Country), .keep = "used")
```

NAs in the input are treated as FALSE, therefore they will return NAs.

```{r}
df %>%
  filter(StockCode == "85166B") %>%
  mutate(Country = na_if(Country, "Unspecified"),
         is_UK = case_when(!Country == "United Kingdom" ~ "No",
                           Country == "United Kingdom" ~ "Yes"), .keep = "used")
```

But we can again use the .default argument to change them to a value of choice.

```{r}
df %>%
  filter(StockCode == "85166B") %>%
  mutate(Country = na_if(Country, "Unspecified"),
         is_UK = case_when(!Country == "United Kingdom" ~ "No",
                           Country == "United Kingdom" ~ "Yes", .default = "not available"), .keep = "used")
```

## - *order of evaluation*

As the tests are evaluated in order, it is important to start from the more particular one and then generalize, to avoid misclassifications when an element can pertain to more than one class, like "Unspecified" here.

```{r}
df %>%
  mutate(is_EU = case_when(Country == "Unspecified" ~ Country,
                           Country %in% EU ~ "Yes",
                           !Country %in% EU ~ "No")) %>% 
  count(is_EU, sort = TRUE)
```

Because, if we had put the "No" condition before the "Unspecified" one, we would have had a column without the "Unspecified" value as "Unspecified" doesn't appear in EU.

```{r}
df %>%
  mutate(is_EU = case_when(Country %in% EU ~ "Yes",
                           !Country %in% EU ~ "No",
                           Country == "Unspecified" ~ Country)) %>% 
  count(is_EU)
```

When present, .default is evaluated last.

As with if_else(), the length of the classes must be consistent across all tests and equal to the number of rows of the data frame or to 1.

It is also important that the length of the logical vectors outputted by the tests are all of the same size (in the second test of the following example that is not the case).

```{r, error = TRUE}
df %>%
  mutate(is_EU = case_when(Country == "Unspecified" ~ Country,
                           EU %in% Country ~ "Yes",
                           !Country %in% EU ~ "No"))
```

All the elements of the new column must be of the same type as well.

```{r, error = TRUE}
df %>%
  mutate(is_EU = case_when(Country == "Unspecified" ~ Country,
                           Country %in% EU ~ TRUE,
                           !Country %in% EU ~ FALSE))
```

As with if_else(), NAs can be evaluated to one of the classes.

```{r}
df %>%
  mutate(Country = na_if(Country, "Unspecified")) %>%
  mutate(is_EU = case_when(Country %in% EU ~ "Yes",
                           !Country %in% EU ~ "No")) %>%
  count(is_EU)
```

If this goes against our wishes, we need to add another test just for them.

```{r}
df %>%
  mutate(Country = na_if(Country, "Unspecified")) %>%
  mutate(is_EU = case_when(is.na(Country) ~ NA,
                           Country %in% EU ~ "Yes",
                           !Country %in% EU ~ "No")) %>%
  count(is_EU)
```

As case_when() first evaluates all of the classes for all of the elements tested, we might get messages that don't seem to apply to our situation, like here where the evaluation of sqrt(Quantity) produces NaNs for the values that are less than 0, NaNs that won't be in the final output though as for those values a 0 will be assigned.

```{r}
df %>%
  mutate(New_Quantity = case_when(Quantity > 0 ~ sqrt(Quantity),
                                  Quantity <= 0 ~ 0))
```

We can use expressions not only for the classes but also for the tests.

```{r}
df %>%
  mutate(New_Quantity = case_when(Quantity / 2 > 0 ~ sqrt(Quantity),
                                  Quantity <= 0 ~ 0))
```

## - *.ptype*

With the .ptype argument we can change the type of the output.

```{r}
df %>%
  mutate(Country = na_if(Country, "Unspecified")) %>%
  mutate(is_EU = case_when(is.na(Country) ~ NA,
                           Country %in% EU ~  TRUE,
                           !Country %in% EU ~ FALSE, .ptype = integer()))
```

## - *.size*

And with the .size one we can reduce the length of the output but I don't understand how to use it.

```{r, error = TRUE}
df %>%
  mutate(Country = na_if(Country, "Unspecified")) %>%
  mutate(is_EU = case_when(is.na(Country) ~ NA,
                           !Country %in% EU ~ FALSE,
                            Country %in% EU ~  TRUE, .size = 1))
```

# - *case_match()*

Thanks to case_match() we can simplify the following example,

```{r}
df %>%
  mutate(Language = case_when(Country %in% c("Austria", "Germany") ~ "Germanic",
                              Country %in% c("France", "Italy", "Portugal", "Spain") ~ "Romance",
                              Country %in% c("Australia", "EIRE", "United Kingdom", "USA") ~ "English",.default = Country), .keep = "used")
```

into this, basically substituting the tests with the values used for the matchmaking.
Notice how the first argument of the function is the column.

```{r}
df %>%
  mutate(Language = case_match(Country,
                               c("Austria", "Germany") ~ "Germanic",
                               c("France", "Italy", "Portugal", "Spain") ~ "Romance",
                               c("Australia", "EIRE", "United Kingdom", "USA") ~ "English", .default = Country), .keep = "used")
```

## - *.ptype*

case_match() has a .ptype argument as well.

```{r}
df %>%
  mutate(English_Language = case_match(Country,
                                       c("Australia", "EIRE", "United Kingdom", "USA") ~ TRUE, .default = FALSE, .ptype = integer()), .keep = "used")
```

## - *with group_by()*

When the data frame is grouped the tests are performed group-wise, so with specific ones the conditions may vary between groups.

```{r}
df %>%
  group_by(Country) %>%
  mutate(Global_Avg_Price = mean(df$Price),
         Local_Avg_Price = mean(Price),
         Exp_Item = case_when(Price > mean(Price) ~ "Expensive", 
                              Price == mean(Price) ~ "Average",
                              Price < mean(Price) ~ "Not Expensive"))
```