---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

`coalesce()` takes more than one vector and,

```{r, df_print = "paged"}
(x <- c("x", "x", NA, "x"))
(y <- c("y", "y", "y", "y"))
```

when the first one specified has NAs elements, they are replaced with elements from the following one that share the same positions.

```{r, df_print = "paged"}
coalesce(x, y)
```

The replacement doesn't work backwards so when there are no NAs to replace in the first vector specified, `coalesce()` just returns it.

```{r, df_print = "paged"}
coalesce(y, x)
```

In case two elements from different vectors are eligible to substitute, `coalesce()` uses the the one from the vector specified first.

```{r, df_print = "paged"}
(z <- c("z", "z", "z", "z"))
coalesce(x, y, z)
coalesce(x, z, y)
```

The second vector can also be a scalar, in this case it will be recycled along the length of the first vector.

```{r, df_print = "paged"}
coalesce(x, "w")
```

# - *usage with a data frame*

In a data frame, we can use this last property to replace the NAs in a column with another value of choice.

```{r}
df %>%
  filter(StockCode == "85166B") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         New_CustomerID = coalesce(`Customer ID`, "missing customer"), .keep = "used")
```

Instead of a single value, we can use several if the length of the vector containing the new values is equal to the size of the column containing the values to be replaced.

```{r}
df %>%
  filter(Invoice == "C489881") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         New_CustomerID = coalesce(`Customer ID`, c("first missing customer", "second missing customer")), .keep = "used")
```

Recycling when length > 1 is not allowed in `dplyr`, even if we use lowest common denominators of the values to be replaced (2 and 8 in the following example), 

```{r, error = TRUE}
df %>%
  filter(StockCode == "85166B") %>%
  select(`Customer ID`)
df %>%
  filter(StockCode == "85166B") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         New_CustomerID = coalesce(`Customer ID`, c("first missing customer", "second missing customer")), .keep = "used")
```

or of the size of the column (3 and 39 in this one).

```{r, error = TRUE}
df %>%
  filter(StockCode == "85166B") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         New_CustomerID = coalesce(`Customer ID`, c("first missing customer", "second missing customer", "third missing costumer")), .keep = "used")
```

We can by all means substitute the missing values of one column with the values of another, as per our introduction to `coalesce()`.

```{r}
df %>%
  filter(StockCode == "85166B") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         For_missing_CustID = "absent",
         New_CustomerID = coalesce(`Customer ID`, For_missing_CustID), .keep = "used")
```

# - *with functions*

`coalesce()` accepts functions as the values to be replaced.

```{r, df_print = "paged"}
pos <- c("first ", "second ")
cust <- "missing customer"
```
```{r}
df %>%
  filter(Invoice == "C489881") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         New_CustomerID = coalesce(`Customer ID`, paste0(pos, cust)), .keep = "used")
```

# - *.ptype*

With the `.ptype` argument we can change the type of the output.

```{r, df_print = "paged"}
coalesce(c(1, NA, 1, 1), 0, .ptype = logical())
```

# - *.size*

And with the `.size` one we can override its length, for example to recycle unitary vectors to a size of choice.

```{r, df_print = "paged"}
coalesce(NA, 1)
coalesce(NA, 1, .size = 6)
```

# - *with group_by()*

A grouped data frame doesn't condition `coalesce()`.

```{r}
df %>%
  group_by(Invoice) %>%
  filter(StockCode == "85166B") %>%
  mutate(`Customer ID` = as.character(`Customer ID`),
         New_CustomerID = coalesce(`Customer ID`, "missing customer"), .keep = "used")
```