---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

`consecutive_id()` is a function that augments a counter every time there is a change in the sequence of elements, in the form of a vector, we feed into it.
The counter starts from 1.

```{r, df_print = "paged"}
consecutive_id(c(1, 2, 1, 3, 4, 3))
```

As long there is a change, the value of the elements is irrelevant.

```{r, df_print = "paged"}
consecutive_id(c(0, 1, 0, 1, 0, 1))
```

NAs are not ignored 

```{r, df_print = "paged"}
consecutive_id(c(0, 1, 0, NA, 1))
```

and, when consecutive, they are treated as the same value.

```{r, df_print = "paged"}
consecutive_id(c(0, 1, 0, NA, NA, 1))
```

# - *usage with a data frame*

`consecutive_id()` comes useful when we want to `group_by()` on distinct sequences made by the same group, like the consecutive purchases made by a customer.

```{r}
df %>%
  mutate(Invoice = Invoice,
         id = consecutive_id(`Customer ID`), .after = "Invoice", .keep = "used")
```

In this way `id` will only augment when the value in the `Customer ID` column changes, so two consecutive invoices, made by the same customer, will have the same `id` (like invoice 489434 and 489435 for customer 13085).

We can then use `id` as a grouping column to calculate for example the total quantity of these consecutive purchases.

```{r}
df %>%
  mutate(id = consecutive_id(`Customer ID`), .after = "Invoice") %>%
  group_by(id, `Customer ID`) %>%
  summarise(Cons_Purch_Total_Quantity = sum(Quantity))
```

If we didn't `group_by(id)` as well, we would have obtained the global total quantity for each customer.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  summarise(Total_Quantity = sum(Quantity))
```

Notice the difference with customer 13085 for example.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  summarise(Total_Quantity = sum(Quantity)) %>%
  filter(`Customer ID` == 13085)
```

With `consecutive_id()` we can isolate consecutive runs of unique values in a column then.

# - *multiple sequences*

`consecutive_id()` accepts multiple sequences as well, as long as they are of the same length.
As they are inspected in parallel and not one after the other, to increase the counter it is sufficient the change in value in one of them.

```{r, df_print = "paged"}
consecutive_id(c(0, 0, 1, 1), 
               c(0, 1, 0, 1))
```

We can take advantage of this feature by imagining that our data frame didn't come with the `Invoice` column, and somehow recreate it by feeding `InvoiceDate` and `Customer ID` to `consecutive_id()`.

```{r}
df %>%
  filter(!is.na(`Customer ID`)) %>%
  mutate(Invoice = Invoice,
         id = consecutive_id(InvoiceDate, `Customer ID`), .after = "Invoice", .keep = "used")
```

As long as every `Invoice` has a different time stamp (and there are no consecutive NAs in the columns processed), `id` will have the same increments as `Invoice`.

```{r}
df %>%
  filter(!is.na(`Customer ID`)) %>%
  mutate(id = consecutive_id(InvoiceDate, `Customer ID`),
         id_Invoice = consecutive_id(Invoice), .after = "Invoice", .keep = "used") %>%
  filter(id != id_Invoice)
```

# - *with group_by()*

On a grouped data frame `consecutive_id()` creates a counter for every group.

```{r}
df %>%
  group_by(Country) %>%
  mutate(id = consecutive_id(Invoice), .after = "Invoice", .keep = "used")
df %>%
  group_by(Country) %>%
  mutate(id = consecutive_id(Invoice), .after = "Invoice") %>%
  distinct(Country, id)
```