---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

consecutive_id() is a function that augments a counter every time there is a change in the sequence of elements, in the form of a vector, we feed to it.
The counter starts from 1.

```{r, df_print = "paged"}
consecutive_id(c(1, 2, 1, 3, 4, 3))
```

As long there is a change, the value of the elements is irrelevant.

```{r, df_print = "paged"}
consecutive_id(c(0, 1, 0, 1, 0, 1))
```

NAs are not ignored and are treated as the same value.

```{r, df_print = "paged"}
consecutive_id(c(0, 1, 0, NA, NA, 1))
```

# - *usage with a data frame*

consecutive_id() comes useful when we want to group_by() on separated sequences made by the same group, like the consecutive purchases made by a customer.

```{r}
df %>%
  mutate(id = consecutive_id(`Customer ID`), .after = "Invoice")
```

In this way id will only augment when the value in the Customer Id column changes, so two consecutive invoices will have the same id.

We can then use id as a grouping variable to calculate for example the total quantity of these consecutive purchases.

```{r}
df %>%
  mutate(id = consecutive_id(`Customer ID`), .after = "Invoice") %>%
  group_by(id, `Customer ID`) %>%
  summarise(Cons_Purch_Total_Quantity = sum(Quantity))
```

If we didn't group_by() id as well, we would have obtained the global total quantity for each customer.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  summarise(Total_Quantity = sum(Quantity))
```

Notice the difference for customer 13085 for example.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  summarise(Total_Quantity = sum(Quantity)) %>%
  filter(`Customer ID` == 13085)
```

With it we can identify consecutive runs of unique values in a column then.

# - *multiple sequences*

consecutive_id() accepts multiple sequences as well, as long as they are of the same length.
As they are inspected in parallel and not one after the other, to increase the counter it is sufficient the change in value in one of them.

```{r, df_print = "paged"}
consecutive_id(c(0, 0, 1, 1), 
               c(0, 1, 0, 1))
```

We can take advantage of this feature by imagining that our data frame didn't come with the Invoice column, and somehow recreating it by feeding InvoiceDate and Customer ID to consecutive_id().

```{r}
df %>%
  filter(!is.na(`Customer ID`)) %>%
  mutate(id = consecutive_id(InvoiceDate, `Customer ID`), .after = "Invoice")
```

As long as every Invoice has a different time stamp (and there are no consecutive NAs in the columns processed), id will have the same increments as Invoice.

```{r}
df %>%
  mutate(id = consecutive_id(InvoiceDate, `Customer ID`),
         id_Invoice = consecutive_id(Invoice), .after = "Invoice") %>%
  filter(id != id_Invoice)
```

# - *with group_by()*

On a grouped data frame consecutive_id() creates a counter for every group.

```{r}
df %>%
  group_by(Country) %>%
  mutate(id = consecutive_id(Invoice), .after = "Invoice")
df %>%
  group_by(Country) %>%
  mutate(id = consecutive_id(Invoice), .after = "Invoice") %>%
  distinct(Country, id)
```