---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr) #OPPURE USO include?
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx")) 
```

data-masking

# - *fundamentals*

`count()` is a wrapper for `group_by() %>% summarise(n = n())` and, like its counterpart, it returns the size, in terms of number of rows, for every group of the data frame.

```{r}
df %>%
  count(Country)
df %>%
  group_by(Country) %>%
  summarise(n = n())
```

As such, it can also be seen as a wrapper for `group_by() %>% tally()` (as `tally()` is a wrapper for `summarise(n = n())`).

```{r}
df %>%
  group_by(Country) %>%
  tally()
```

But they differ when there is more than one grouping column, as `tally()`, like `summarise()`, removes only the most recent grouping,

```{r, df_print = "paged"}
df %>%
  group_by(Country, `Customer ID`) %>%
  tally() %>%
  group_vars()
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(n = n()) %>%
  group_vars()
```

while `count()` ungroups the output even with several columns, saving us from piping an additional `ungroup()`.

```{r, df_print = "paged"}
df %>%
  count(Country, `Customer ID`) %>%
  group_vars()
```

When we specify more columns, `count()` returns the number of rows for all the unique existing combinations between their values.

```{r}
df %>%
  count(Country, `Customer ID`)
```

Like if we specified two columns in `group_by()`.

```{r}
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(n = n())
```

As with `group_by() %>% summarise(n = n())`, NAs are counted as one value.

```{r}
df %>%
  count(`Customer ID`, sort = TRUE)
df %>%
  group_by(`Customer ID`) %>% 
  summarise(n = n()) %>%
  arrange(desc(n))
```

So with more than one column we can have every existing combination of NAs with other values.

```{r}
df %>%
  count(`Customer ID`, Description, sort = TRUE)
```

Like if we wrote

```{r}
df %>%
  group_by(`Customer ID`, Description) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```

Being a data-masking function, we can use it with expressions. 

```{r}
df %>%
  count(CustomerID = `Customer ID`)
df %>%
  count(Price_Rank = dense_rank(Price))
```

This is equivalent to using those expressions in the `group_by()` calls.

```{r}
df %>%
  group_by(CustomerID = `Customer ID`) %>%
  summarise(n = n())
df %>%
  group_by(Price_Rank = dense_rank(Price)) %>%
  summarise(n = n())
```

# - *sort*

The `sort` argument can be employed if we want the rows with the largest `n` on top, without having to pipe an `arrange()` call afterwards.

```{r}
df %>%
  count(Country, sort = TRUE)
df %>% 
  count(Country, `Customer ID`, sort = TRUE)
```

# - *wt*

data-masking

Another useful argument is `wt` (shorthand for "weight"), which allows us to rank one column by the values of another, like here where we return for each country the summation of their values in the `Quantity` column and not just how many rows they have.

```{r}
df %>%
  count(Country, wt = Quantity)
```

Internally it modifies the `summarise()` call from `n()` to `sum(wt)`.

```{r}
df %>%
  group_by(Country) %>%
  summarise(Total_Quantity = sum(Quantity))
```

So, instead of counting each row as 1, it counts each row by the values of the column used in `wt`.

Using columns with values that can't be summed results with an error.

```{r, error = TRUE}
df %>%
  count(Country, wt = Description)
```

If our data has already been aggregated, 

```{r}
df %>%
  count(Country, `Customer ID`)
```

`wt` can be employed to retrieve the original number of rows pertaining to a group.

```{r}
df %>%
  count(Country, `Customer ID`) %>%
  count(Country, wt = n)
```

`wt` also offers the possibility of being used with a function.

```{r}
df %>%
  count(Country, wt = range(Quantity))
```

That example translates to this,

```{r}
df %>%
  group_by(Country) %>%
  summarise(n = sum(range(Quantity)))
```

where it's easier to understand that, for every value of the `Country` column, we sum the two values returned by `range(Quantity)`.

```{r}
df %>%
  group_by(Country) %>%
  reframe(n = range(Quantity))
```

We can also use functions that only return one value per group,

```{r}
df %>%
  count(Country, wt = n_distinct(Invoice))
```

but in this case the summation is inconsequential as there is only one addend

```{r}
df %>%
  group_by(Country) %>%
  summarise(n = sum(n_distinct(Invoice)))
df %>%
  group_by(Country) %>%
  summarise(n = n_distinct(Invoice))
```

and it is mainly just a way to use another function beside `n()` in the, wrapped, `summarise()` call of `count()`.

Using a function that doesn't output a number clearly throws an error.

```{r, error = TRUE}
df %>%
  count(Country, wt = stringr:::str_replace(`Customer ID`, "2", "3"))
```

`wt` can be used with more than one column as well.

```{r}
df %>%
  count(Country, wt = c(Quantity, Price))
```

And it would be like summing together, for every value of the grouping column, every value of the columns specified.

```{r}
df %>%
  group_by(Country) %>%
  summarise(n = sum(Quantity, Price))
```

Another example, here we return 1 (TRUE) for the countries that have more than 15 rows and 0 (FALSE) for the ones that don't (only Lebanon here).

```{r}
df %>%
  count(Country, wt = n() > 15)
```

As in one of the previous examples, the summation is inconsequential, apart from transforming logical values into integers.

```{r}
df %>%
  group_by(Country) %>%
  summarise(n = sum(n() > 15))
df %>%
  group_by(Country) %>%
  summarise(n = n() > 15)
```

# - *name*

The `name` argument, which modifies the name of the `count()` output column from the default `n`,

```{r}
df %>%
  count(Country, name = "N_Rows")
```

comes in handy with `wt` to have more apt denominations, like we can do with `summarise()`.

```{r}
df %>%
  count(Country, wt = Quantity, name = "Total_Quantity")
df %>%
  group_by(Country) %>%
  summarize(Total_Quantity = sum(Quantity))
```

Using more than one column is not a problem.

```{r}
df %>%
  count(Country, `Customer ID`, wt = n_distinct(Invoice), name = "N_Invoices")
```

# - *.drop*

The last argument is `.drop`, which, when we set to FALSE (default is TRUE), has the same utility as in `group_by()`, alas it doesn't drop the levels of a factor when their count is equal to 0 (the Unspecified country in this example, that has 0 rows after we modify its name to NA if at the same time we retain it as a level).

```{r}
df %>%
  mutate(Country = na_if(Country, "Unspecified"),
         Country = factor(Country, levels = unique(df$Country))) %>%
  count(Country, .drop = FALSE)
df %>%
  mutate(Country = na_if(Country, "Unspecified"),
         Country = factor(Country, levels = unique(df$Country))) %>%
  count(Country)
```

`.drop = FALSE` will keep the empty levels also when they have already been removed by a previous manipulation (`filter(n() > 15)` here, which removes the Lebanon group).

```{r}
df %>%
  group_by(Country = factor(Country)) %>%
  filter(n() > 15) %>%
  count(Country, .drop = FALSE)
df %>%
  group_by(Country = factor(Country)) %>%
  filter(n() > 15) %>%
  count(Country)
```

As with `group_by()`, we must be careful when using several factor columns as it returns combinations not present in the data frame, like every customer located in every country.

```{r}
df %>%
  count(factor(Country), factor(`Customer ID`), .drop = FALSE)
df %>%
  count(factor(Country), factor(`Customer ID`))
```

# - *repetitions spotter*

As already seen, with more than one column `count()` returns all the unique existing combinations between their values, counting with `n` the number of rows pertaining to each group.

```{r}
df %>%
  count(Invoice, StockCode)
```

If `n` is higher than 1, like for invoice 489488 with stock code 22125,

```{r}
df %>%
  count(Invoice, StockCode) %>%
  filter(Invoice == "489488")
```

it means that there are several rows with the same values in, at least, those two columns.

```{r}
df %>%
  filter(Invoice == "489488" &
           StockCode == "22125")
```

Filtering for `n > 1` can then be used as a way to spot repetitions in the columns specified in `count()`, which can be errors if those two columns should have had a single unique combination.

```{r}
df %>%
  count(Invoice, StockCode) %>%
  filter(n > 1)
```

If desired, we can extend this procedure to all the columns to spot duplicated rows.

```{r}
df %>%
  count(pick(everything()), name = "# of repetitions") %>%
  filter(`# of repetitions` > 1)
```

If instead we are interested in spotting the non duplicated rows, we can just use `distinct()`.

```{r}
df %>%
  distinct()
```

# - *number of unique combinations*

The procedure can be extended if we want to know how many unique existing combinations of two columns there are for each value of one of them.
In the following example the one column is `Invoice` and we are looking for the unique existing combinations with `StockCode` for every value of it.

```{r}
df %>%
  count(Invoice, StockCode) %>%
  count(Invoice)
```

So for instance for invoice 489488 (the invoice from a previous example) there are 29 unique combinations with the `StockCode` column.

```{r}
df %>%
  count(Invoice, StockCode) %>%
  count(Invoice) %>%
  filter(Invoice == "489488")
```

That obviously amounts to the number of rows from the first `count()` call.

```{r}
df %>%
  count(Invoice, StockCode) %>%
  filter(Invoice == "489488")
```

If we add a third column as well (to look for the unique existing combinations of three columns for every value of one of them), `Price` for example, `n` stays the same,

```{r}
df %>%
  count(Invoice, StockCode, Price) %>%
  count(Invoice) %>%
  filter(Invoice == "489488")
```

as `Price` doesn't vary for any stock code, so the number of combinations is constant.

```{r}
df %>%
  count(Invoice, StockCode, Price) %>%
  filter(Invoice == "489488") %>%
  arrange(StockCode)
```

This can more easily checked with a grouped `filter()` on the number of rows per stock code: if no stock code has more than one row per invoice, it means that only one value of `Price` is assigned to it.

```{r}
df %>%
  count(Invoice, StockCode, Price) %>%
  filter(Invoice == "489488") %>%
  group_by(StockCode) %>%
  filter(n() > 1)
```

If we use `Quantity`instead of `Price` though, `n` increases of one unit,

```{r}
df %>%
  count(Invoice, StockCode, Quantity) %>%
  count(Invoice) %>%
  filter(Invoice == "489488")
```

because that column changes values between two rows for stock code 22125, as we've seen in a previous example, resulting therefore in an additional combination,

```{r}
df %>%
  count(Invoice, StockCode, Quantity) %>%
  filter(Invoice == "489488") %>%
  arrange(StockCode)
```

that, as before, can be more easily spotted when we group filter for `n()` > 1 on stock codes.

```{r}
df %>%
  count(Invoice, StockCode, Quantity) %>%
  filter(Invoice == "489488") %>%
  group_by(StockCode) %>%
  filter(n() > 1)
```

If we wish to generalize the grouped `filter()` call, we need to add the `Invoice` column as a grouping one,

```{r}
df %>%
  count(Invoice, StockCode, Quantity) %>%
  group_by(Invoice, StockCode) %>%
  filter(n() > 1)
```

otherwise we would just filter on stock codes regardless of the invoice they pertain to.

```{r}
df %>%
  count(Invoice, StockCode, Quantity) %>%
  group_by(StockCode) %>%
  filter(n() > 1) %>%
  arrange(StockCode)
```

Instead of piping two `count()`s in succession we can also use the `wt` argument with `n_distinct()` to attain the same results.

```{r}
df %>%
  count(Invoice, StockCode) %>%
  count(Invoice)
df %>%
  count(Invoice, wt = n_distinct(StockCode))
```

That translates to

```{r}
df %>%
  group_by(Invoice) %>%
  summarise(n = n_distinct(StockCode))
```

We can as well use more than one columns with `n_distinct()`, always to return the number of all the unique existing combinations of three columns for every value of one of them.

```{r}
df %>%
  count(Invoice, StockCode, Quantity) %>%
  count(Invoice) %>%
  filter(Invoice == "489488")
df %>%
  count(Invoice, wt = n_distinct(StockCode, Quantity)) %>%
  filter(Invoice == "489488")
```

So we need to use the one column in `count()` and the other two in `n_distinct()`.

Without `count()` we would otherwise write like this.

```{r}
df %>%
  group_by(Invoice) %>%
  summarise(n = n_distinct(StockCode, Quantity)) %>%
  filter(Invoice == "489488")
```

But if we instead write like this, with only one column in `n_distinct()`,

```{r}
df %>%
  count(Invoice, StockCode, wt = n_distinct(Quantity)) %>% 
  filter(Invoice == "489488")
```

that translates to this,

```{r}
df %>%
  group_by(Invoice, StockCode) %>%
  summarise(n = n_distinct(Quantity)) %>%
  filter(Invoice == "489488")
```

we would obtain the number of unique values of the `Quantity` column for every unique existing combination of `Invoice` and `StockCode` (something very different from before then, as shown by the output).

As before, a subsequent grouped filter on `n` can be useful to more easily pinpoint the stock codes with different quantities in the same invoice.

```{r}
df %>%
  count(Invoice, StockCode, wt = n_distinct(Quantity)) %>% 
  group_by(Invoice, StockCode) %>%
  filter(n > 1)
```

# - *with group_by()*

When used on a grouped data frame, `count()` appends the groups created by its argument, performs the count and then ungroups while retaining the original grouping.

```{r}
df %>%
  group_by(Invoice) %>%
  count(StockCode, Quantity)
```

It is equivalent to

```{r}
df %>%
  group_by(Invoice) %>%
  group_by(StockCode, Quantity, .add = TRUE) %>%
  summarise(n = n()) %>%
  ungroup(StockCode, Quantity)
```

If we add the column from `group_by()` to `count()`, as a way to modify the columns' order of the output, it is important to notice that `count()` doesn't overwrite it, as the column from `group_by()` has the precedence, being the group key of the output.

```{r}
df %>%
  group_by(Invoice) %>%
  count(StockCode, Invoice, Quantity)
```

So we have to resort to pipe an additional `select()` call.

```{r}
df %>%
  group_by(Invoice) %>%
  count(StockCode, Quantity) %>%
  select(StockCode, Invoice, Quantity, n)
```

With `sort` we can rearrange the rows though, as it overrides the group keys from `group_by()`.

```{r}
df %>%
  group_by(Invoice) %>%
  count(StockCode, Quantity, sort = TRUE)
```

# - *add_count()*

data-masking

In case we don't want to summarize our data frame, we can use `add_count()` which is equivalent to `group_by() %>% mutate(n = n()`).

```{r}
df %>%
  add_count(Country)
df %>%
  group_by(Country) %>% 
  mutate(n = n())
```

Like `count()`, it ungroups the output also with several columns,

```{r}
df %>%
  add_count(Country,`Customer ID`)
```

while `mutate()` keeps the grouping as is.

```{r}
df %>%
  group_by(Country, `Customer ID`) %>% 
  mutate(n = n())
```

`add_count()` can be used with expressions

```{r}
df %>%
  add_count(CustomerID = `Customer ID`)
df %>%
  add_count(Price_Rank = dense_rank(Price))
```

and is consistent with the optional arguments of `count()`,

```{r}
df %>%
  add_count(Country, wt = Quantity, name = "Total_Quantity", sort = TRUE)
```

except for `.drop`, deprecated, as it can't modify the output.

```{r}
df %>%
  mutate(Country = na_if(Country, "Unspecified"),
         Country = factor(Country, levels = unique(df$Country))) %>%
  add_count(Country, .drop = FALSE)
df %>%
  mutate(Country = na_if(Country, "Unspecified"),
         Country = factor(Country, levels = unique(df$Country))) %>%
  add_count(Country)
```

For instance in the following example it can't reconstruct previously filtered rows.

```{r}
df %>%
  group_by(Country = factor(Country)) %>%
  filter(n() > 15) %>%
  add_count(Country, .drop = FALSE)
```

## - *with group_by()*

`add_count()` on a grouped data frame has the same behavior as `count()` (appending the groups created by its argument, performing the count and then ungrouping while keeping the original grouping).

```{r}
df %>%
  group_by(Invoice) %>%
  add_count(StockCode, Quantity)
```

Similarly as before, it is equivalent to this.

```{r}
df %>%
  group_by(Invoice) %>%
  group_by(StockCode, Quantity, .add = TRUE) %>%
  mutate(n = n()) %>%
  ungroup(StockCode, Quantity)
```

Likewise it can't modify the columns' order of the output,

```{r}
df %>%
  group_by(Invoice) %>%
  add_count(StockCode, Invoice, Quantity)
```

but `sort` can modify the rows' order.

```{r}
df %>%
  group_by(Invoice) %>%
  add_count(StockCode, Quantity, sort = TRUE)
```