---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

data-masking

# - *fundamentals*

When there are rows that are repeated,

```{r}
df %>%
  filter(Invoice == "496431" & 
           StockCode == "84826")
```

`distinct()` removes the ones after the first.

```{r}
df %>%
  filter(Invoice == "496431" & 
           StockCode == "84826") %>%
  distinct()
```

So it is useful in data wrangling when we want to remove duplicated rows from a data frame.

```{r}
df %>%
  distinct()
```

It is equivalent to these lines of code,

```{r}
df %>%
  group_by(pick(everything())) %>%
  summarise(n = n()) %>%
  select(-n) %>%
  ungroup()
df %>%
  group_by(pick(everything())) %>%
  slice(1) %>%
  ungroup()
```

but it returns the output much faster.

If we are interested in returning the duplicated rows instead, we can use this procedure with `count()` and `filter()`.

```{r}
df %>%
  count(pick(everything()), name = "# of repetitions") %>%
  filter(`# of repetitions` > 1)
```

We can also use it on a subset of columns and it will return the unique values, in case of one, and the existing combinations of their values with more than one.

```{r}
df %>%
  distinct(Country)
df %>%
  distinct(Country, `Customer ID`)
```

# - *.keep_all*

With `.keep_all = TRUE`, we keep all other columns as well in the output.

```{r}
df %>%
  distinct(Country, .keep_all = TRUE)
df %>%
  distinct(Country, `Customer ID`, .keep_all = TRUE)
```

If we don't specify columns, the outputs are equal.

```{r}
df %>%
  distinct()
df %>%
  distinct(.keep_all = TRUE)
```

`distinct()` is very similar, beside the added count and the row order of the output, to what `count()` does.

```{r}
df %>%
  count(Country)
df %>%
  count(Country, `Customer ID`)
```

NAs are treated as one value.

```{r}
df %>%
  distinct(`Customer ID`) %>%
  arrange(!is.na(`Customer ID`))
```

So with more than one column they form a combination with the other values.

```{r}
df %>%
  mutate(Country = na_if(Country, "Unspecified")) %>%
  distinct(Country, `Customer ID`) %>%
  arrange(!is.na(Country),!is.na(`Customer ID`))
```

We can use `pick()` to simplify the selection of more than one column.

```{r}
df %>%
  distinct(pick(starts_with("I")))
```

It being a data-masking function, we can also use expressions. 

```{r}
df %>%
  distinct(Invoice_Day = as.Date(InvoiceDate))
```

`distinct()` only works with data frames (a single column one included) but not with vectors.

```{r, error = TRUE, df_print = "paged"}
df %>%
  pull(Country) %>%
  distinct()
distinct(df$Country)
```

So in cases like these we have to rely on `unique()` from base `R`.

```{r, df_print = "paged"}
df %>%
  pull(Country) %>%
  unique()
```

# - *with group_by()*

With a grouped data frame, the grouping column is processed as well, as if it was specified first.
The rows order is kept from `df` and not changed following the grouping columns.

```{r}
df %>%
  group_by(Country) %>%
  distinct(`Customer ID`)
```