---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

We will here discuss several functions that are useful with grouped data frames.

# - *group_split()*

This function divides the groups of a grouped data frame into separate tibbles.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_split()
df %>%
  slice(1:100) %>%
  group_by(Country, `Customer ID`) %>%
  group_split()
```

The output is a list, arranged by the grouping columns, so every data frame can be accessed with the correct subsetting (based on indexes as the list is not named.)

```{r}
dfs_list <- df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_split()
dfs_list[[1]]
```

We can modify that by using `setNames()` though.

```{r, df_print = "paged"}
(named_dfs_list <- df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_split() %>%
  setNames(unique(sort(df[1:100, ]$Country))))
```
```{r}
named_dfs_list[["France"]]
```

If we want to merge the data frames back, we can use `bind_rows()`.

```{r}
bind_rows(dfs_list)
```

The rows' order might be different from the original data frame's one as it is dictated by the grouping columns.

The function can be used directly without having to `group_by()` first,

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_split(Country)
```

also with expressions.

```{r, df_print = "paged"}
df %>%
  slice(1:10) %>%
  group_split(Price_Rank = dense_rank(Price))
```

## - *.keep*

There is a `.keep` argument that controls whether to keep or not the grouping columns inside the tables, the default is TRUE.

```{r}
dfs_list_keep <- df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_split(.keep = FALSE)
dfs_list_keep[[1]]
```

# - *group_nest()*

`group_nest()` condenses a grouped data frame into a list column with one row for every combination of the grouping columns.

```{r}
df %>%
  group_by(Country) %>%
  group_nest()
df %>%
  group_by(Country, `Customer ID`) %>%
  group_nest()
```

Every tibble of the list column contains the rows relative to that combination of the grouping columns.

```{r}
df %>%
  group_by(Country) %>%
  group_nest() %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
df %>%
  group_by(Country, `Customer ID`) %>%
  group_nest() %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

It can be used directly on an ungrouped data frame as well.

```{r}
df %>%
  group_nest(Country)
```

We can unnest to the original data frame with the relative function from `tidyr`, but the rows' order will be different, following the one of grouping columns.

```{r}
df %>%
  group_by(Country) %>%
  group_nest() %>%
  tidyr:::unnest(cols = c(data))
```

We can nest by expressions.

```{r}
df %>%
  slice(1:100) %>%
  group_nest(Price_Rank = dense_rank(Price))
```

## - *.key*

The function has two optional arguments: with `.key` we can change the name of the list column.

```{r}
df %>%
  group_by(Country) %>%
  group_nest(.key = "list column")
```

## - *keep*

And with `keep` we decide whether or not to keep the grouping columns in the list column's tibbles (the default is FALSE, not keeping them).

```{r}
df %>%
  group_by(Country) %>%
  group_nest() %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
df %>%
  group_by(Country) %>%
  group_nest(keep = TRUE) %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

Using `unnest()` on all the list column while `keep = TRUE` returns an error.

```{r, error = TRUE}
df %>%
  group_by(Country) %>%
  group_nest(keep = TRUE) %>%
  tidyr::unnest(cols = c(data))
```

# - *group_trim()*

`group_trim()` suppresses the levels of the factors used to `group_by()`, in case they are empty after a manipulation (in this case Lebanon).

```{r}
df %>%
  group_by(factor(Country)) %>%
  filter(n() > 15) %>%
  group_trim() %>%
  group_keys()
```

# - *group_map()/modify()/walk()*

With `group_map()` we can apply a function, with a formula syntax, to every group.
The result is a list.

So if for example we want the average quantity for every country,

```{r}
df %>%
  group_by(Country) %>%
  summarise(Avg_Quantity = mean(Quantity))
```

we can write:

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  group_map(~ mean(.x$Quantity))
```

`group_modify()` instead returns a data frame, so we have to use it with functions whose output is of that class.

```{r}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_modify(~ rows_insert(.x, tibble(Invoice = "xxxxxx"), by = "Invoice"))
```

With `group_walk()` the function we apply will only be evaluated for its possible side effects and it will always return the input `.x` unaffected.

```{r}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_walk(~ rows_insert(.x, tibble(Invoice = "xxxxxx"), by = "Invoice"))
```

`group_walk()` can then be used, for example, to print plots (as that is a side effect of `plot()`) in the middle of a pipe.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_walk(~ plot(.x$Quantity)) %>%
  group_map(~ mean(.x$Quantity))
```

In all of these past examples `.x` referred to the object, that is the grouped data frame, that was piped into one of these three `group_*` functions but we can use `.y` as well when we want to refer to the grouping columns.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_map(~ paste(.x$`Customer ID`, .y))
```

All of these functions can also be used on ungrouped data frames,

```{r, df_print = "paged"}
df %>%
  group_map(~ mean(.x$Quantity))
```

with a custom function (that needs two arguments, hence the `...`, even if they don't do nothing in this instance),

```{r, df_print = "paged"}
custfun <- function(x, ...){
  head(x)
}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_map(custfun)
```

and they all have a `.keep` argument, to preserve the grouping columns in the output.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_map(custfun, .keep = TRUE)
```

# - *group_data()*

`group_data()` stores, for each group, every row index pertaining to that group in a list column named `.rows`.

```{r}
df %>%
  group_by(Country) %>%
  group_data()
df %>%
  group_by(Country) %>%
  group_data() %>%
  tidyr::unnest(cols = c(.rows))
```

# - *group_rows()*

`group_rows()` returns the rows' indexes of every group as a list of ordered vectors.

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  filter(Country %in% c("Australia", "Lebanon")) %>%
  group_rows()
```

# - *group_keys()*

`group_keys()` returns a data frame with all the unique values of the grouping columns, if one,

```{r}
df %>%
  group_by(Country) %>%
  group_keys()
```

and one with as many columns as the ones specified as arguments of `group_by()`, if several.

```{r}
df %>%
  group_by(Country, `Customer ID`) %>%
  group_keys()
```

In this case the number of rows depends on the number of unique existing combinations between the grouping columns.

```{r}
df %>%
  group_by(StockCode, Description) %>%
  group_keys()
df %>%
  group_by(StockCode) %>%
  slice(1) %>%
  group_by(StockCode, Description) %>%
  group_keys()
```

The `group_keys()` are also what functions like `summarise()` and `slice(`) order their output by.

```{r}
df %>%
  group_by(Country) %>%
  summarise(Avg_Price = mean(Price))
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(Avg_Price = mean(Price))
df %>%
  group_by(Country, `Customer ID`) %>%
  slice(1)
```

# - *group_indices()*

`group_indices()` returns a vector with a group identifier, in form of a digit, for every row of the data frame.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_by(Country) %>%
  group_indices()
```

The indices are not assigned randomly, but following the order of the `group_keys()`.

# - *group_vars()*

`group_vars()` lists the grouping columns as a vector.

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  group_vars()
df %>%
  group_by(Country, `Customer ID`) %>%
  group_vars()
```

# - *groups()*

`groups()` does the same, but as a list.

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  groups()
df %>%
  group_by(Country, `Customer ID`) %>%
  groups()
```

# - *group_size()*

`group_size()` supplies the size (as per the number of rows) of each group as a unnamed vector.

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  group_size()
```

As always, the order is dictated by the `group_keys()`.

# - *n_groups()*

`n_groups()` supplies the overall number of groups.

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  n_groups()
```

# - *context*

This family of functions, to be used inside a `mutate()` or `summarise()` / `reframe()` call, provides information about the groups.

## - *n()*

`n()` returns the number of rows per group as a data frame.

```{r}
df %>%
  group_by(Country) %>%
  summarise(n = n())
```

NAs are counted as one value.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```

`n()` can be used in other functions as well,

```{r}
df %>%
  filter(Country == "Lebanon") %>%
  slice(n())
```

as a replacement of `nrow(df)` for when we altered the original data frame 

```{r}
df %>%
  filter(Country == "Lebanon") %>%
  slice(nrow(df))
```

and we don't want to use the `magrittr` dot placeholder.

```{r}
df %>%
  filter(Country == "Lebanon") %>%
  slice(nrow(.))
```

Here the data frame was not grouped, so `n()` provided the total number of rows, as if there was just one group.

## - *cur_group()*

`cur_group()` returns one tibble for every combination of the grouping variables, with every one of them containing the relative `group_keys()`.

```{r}
df %>%
  group_by(Country) %>%
  summarise(data = cur_group())
df %>%
  group_by(Country) %>%
  summarise(data = cur_group()) %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(data = cur_group())
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(data = cur_group()) %>%
  ungroup() %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

We can use `cur_group()` to apply manipulations that change depending on the group they are applied to, like filtering on only one group.

```{r}
df %>%
  group_by(Country) %>%
  filter(cur_group()$Country != "United Kingdom" | Quantity > 10)
```

In this example the first test checks, for every group, on whether the value of the `Country` column in the tibble provided by `cur_group()` is different from United Kingdom.

In case of TRUE, we will preserve every row, regardless of the second test, as we specified an OR condition.

In case of FALSE, we will preserve only the rows where the second test is TRUE, which are the ones where the `Quantity > 10` and the country is United Kingdom. 

We can also filter each group for a different value.

These different values must be stored in a tibble, and the correct one for each group will be accessed by subsetting it with the grouping column of the tibble returned by `cur_group()`.

```{r, df_print = "paged"}
Min_Quantity <- tibble("France" = 2,
                       "Italy" = 10,
                       "Germany" = 5)
```
```{r}
df %>%
  filter(Country %in% c("France", "Italy", "Germany")) %>%
  group_by(Country) %>%
  filter(Quantity >= Min_Quantity[[cur_group()$Country]])
```

## - *cur_group_id()*

`cur_group_id()` supplies a numeric identifier for every group, assigned following the order of the `group_keys()`.

```{r}
df %>%
  group_by(Country) %>%
  mutate(id = cur_group_id())
```

## - *cur_group_rows()*

`cur_group_rows()` returns the rows indices for every row of a grouped data frame, sorted by the `group_keys()`.

```{r}
df %>%
  group_by(Country) %>%
  reframe(row = cur_group_rows())
df %>%
  group_by(Country, `Customer ID`) %>%
  reframe(row = cur_group_rows())
```