---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

With `if_else()` we can classify the elements of a vector

```{r, df_print = "paged"}
(x <- c(1, 2, 2, 3))
```

by testing a condition on each one of them.

```{r, df_print = "paged"}
x == 1
```

The first argument is the test, then we need to input the classes that satisfy the TRUE and FALSE outputs.

```{r, df_print = "paged"}
if_else(x == 1, "1", "not1")
```

The type of these two classes must be the same.

```{r, error = TRUE, df_print = "paged"}
if_else(x == 1, 1, "not1")
```

If an element of the vector doesn't satisfy neither class, 

```{r, df_print = "paged"}
(xNA <- c(x, NA))
xNA == 1
```

the function returns NA.

```{r, df_print = "paged"}
if_else(xNA == 1, "1", "not1")
```

# - *missing*

We can change that with the `missing` argument thought, whose value must be of the same type of the other two classes.

```{r, df_print = "paged"}
if_else(xNA == 1, "1", "not1", missing = "other")
```

It is important to pay attention to NAs, as they can be evaluated to one of the two classes.

```{r, df_print = "paged"}
(y <- c(1, 8, 9))
xNA %in% y
if_else(xNA %in% y, TRUE, FALSE)
```

If we want to keep them as they are we can nest two `if_else()`s and use `is.na()`.

```{r, df_print = "paged"}
if_else(is.na(xNA), NA, 
        if_else(xNA %in% y, TRUE, FALSE))
```

Even when nested, the classes must be of the same type.

```{r, error = TRUE, df_print = "paged"}
if_else(is.na(xNA), TRUE, 
        if_else(xNA %in% y, "Yes", "No"))
```

The classes can be of the same length of the output of the test, not solely of length 1.

```{r, df_print = "paged"}
rank_1 <- c("first_is_1", "second_is_1", "third_is_1", "fourth_is_1")
rank_not1 <- c("first_is_not1", "second_is_not1", "third_is_not1", "fourth_is_not1")
if_else(x == 1, rank_1, rank_not1)
```

# - *ptype*

With the `ptype` argument we can change the type of the output.

```{r, df_print = "paged"}
if_else(x == 1, TRUE, FALSE, ptype = integer())
```

# - *size*

And with the `size` one its length, but I don't understand how it is supposed to function.

```{r, error = TRUE, df_print = "paged"}
if_else(x == 1, TRUE, FALSE, size = 8)
if_else(x[1] == 1, TRUE, FALSE, size = 4)
```

# - *usage with a data frame*

With a data frame, `if_else()` is useful to create new columns.

```{r, df_print = "paged"}
EU <-  c("Austria", "Belgium", "Channel Islands", "Cyprus", "Denmark", "EIRE", "Finland", "France", "Germany", 
         "Greece", "Iceland", "Italy", "Lithuania", "Malta", "Netherlands", "Norway", "Poland", "Portugal", 
         "Spain", "Sweden", "Switzerland")
```
```{r}
df %>%
  mutate(in_EU = if_else(Country %in% EU, TRUE, FALSE), .keep = "used")
```

It is important that the output of `if_else()` is either of length 1 or of the same size as the number of rows of the data frame, otherwise we would get an error.

```{r, error = TRUE}
df %>%
  mutate(is_EU = if_else(EU %in% Country, TRUE, FALSE))
```

The test can be specified with expressions.

```{r}
df %>%
  mutate(High_Rank_Price = if_else(min_rank(Price) < 4, TRUE, FALSE), .keep = "used")
```

# - *with group_by()*

When the data frame is grouped the tests are performed group-wise, so with specific ones the output may vary between groups.

For instance a `Price` of 4.25 is in the following example labeled as "Expensive" for Germany (row 32) but not for United Kingdom.

```{r}
df %>%
  group_by(Country) %>%
  mutate(Local_Avg_Price = mean(Price),
         Exp_Item = if_else(Price > mean(Price), "Expensive", "Not Expensive"), .keep = "used") %>%
  filter(between(Price, 4.189418, 4.543470))
```