---
title: "disclaimer"
author: ""
date: ""
output: 
  html_document
---

THIS SITE IS STILL A WORK IN PROGRESS

# - *MOTIVATION*

This site started after the release of dplyr 1.0.0 in 2020, in the form of a Rmarkdown where I noted down the changes from previous versions.

During the years that Rmarkdown grow bigger and bigger, both in size and in scope, not being anymore a changelog of updates but also containing the basics of various functions.

At the end of 2022 I decided to restructure the file, to give it more utility and also to ease its consultation. The end result of that process is this site, as along the weeks I decided to be very throughout and detailed about all the functions that exist in dplyr.

It was both a way to challenge me in producing helpful technical resources in a language that is not my mother tongue, and to further my knowledge of dplyr and of R at a large, as I feel that is the main tool for a Data Analyst like myself.

This site went through many revisions and verification, still some mistakes can always occur. I apologize if those will cause troubles, or if my exposition is not the best. 

# - *INTENDED AUDIENCE*

The intended audience was... myself, as I planned to use this blog as a personal reference guide when I have doubts.

Writing it though I was as well imagining a reader that has little knowledge about dplyr (and Data Analysis) as a whole and wants to deepen their acquaintance with the package.

It is my way to give back to the community, to which I am extremely grateful as it was thanks to others people teaching that I could land my first job in the Data Analysis field.

This site is not a course though, as in every page I may use concepts explained in some other pages so for somebody that wants to learn dplyr from 0 I feel that this would not be the best place to start.

# - *STYLE*

With that intended audience in mind, I strove to be as clear and intelligible as possible, through the following means:  
- one information/concept at a time  
- an easy example after every information/concept provided  
- using the same variables in different examples   
- common language instead of technical one, like I always used column and row instead of variable and observation, never x and y when referring to objects and so on and so forth.

About the site's visual presentation, I kept it simple and clean, also because I didn't nor don't want to spend too much time on it.

# - *MISSING TOPICS*

joining functions
remote tables verbs https://dplyr.tidyverse.org/reference/index.html#remote-tables

# - *THE DATA FRAME*

I always used the same data frame for all the examples provided, as I think that keeping that as a constant reduce the cognitive load when learning new concepts.

The data frame I used can be found here:
https://archive.ics.uci.edu/ml/datasets/Online+Retail+II

and I chose it primarily for a personal convenience as I was already using it for another project of mine. METTERE LINK

# - *SHORTCOMINGS*

IMPAGINAZIONE

mettere virgolette attorno a tutte le funzioni e colonne di df, anche per pkgs 
modificare le liste come ho fatto per across.html?


ultimo controllo certosino
filter()
relocate()/rename() when grouped

stackoverflow
numero dei gruppi NON PER rowwise()/nest_by() -> stackoverflow


-in filter/group_by/utility_functions/across_pick viene troppo lungo e non si riesce a scrollare fino in fondo -> menu superiore a scomparsa? (_site.yml) toc_float: true?

```{r}
# df %>%
#   filter(StockCode == "90081B") %>% #per questo in slice_max non funziona la freccina
#   slice_max(`Customer ID`, n = 3)
```
```{r}
# df %>%
#   count(across(everything())) %>% #COME MAI NON MI SUGGERISCE pick() QUI?
#   filter(n > 1)
```


ATTENZIONE CHE <DATA-MASKING> NON VIENE RENDERATO, CONTROLLARE

R-MARKDOWN
poter copiare il codice con un click (klippy)
aumentare dimensione df?
print() carino in mezzo a pipe
spazio bianco in fondo
un solo readxl (ho provato con cache ma non funziona) e info_paged per tutti i markdown?

CONTENUTI (+memorandum, interesting, etc. etc.)
-arrange esempio in group_by bisogna scrollare le colonne per vederlo?
-group_by ricontrollo implicit and explicit, 553 dopo count()
-coalesce deve essere stessa classe? problema in dubbi.R
-bind aggiungere names repair in bind_cols

GIT
capire Force Push
devo fare Build in RStudio ogni volta che aggiorno?

# - *ACKOWNLEDGEMENTS*

shafee stackoverflow

# - *CONTACTS INFORMATION*

# - *LICENSE OF USE/TERMS OF AGREEMENT*

DEVO SCEGLIERE LA LICENZA, ANCHE PER LE REPO SU GITHUB 

# - *LAST UPDATE & FUTURE ENDEVOURS*

This site was last updated on the xx of xxx 2023, using the x.x.x version of dplyr