---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

`na_if()` substitutes a specified value of a column with NAs.

```{r}
df %>%
  filter(StockCode == "17011F") %>%
  mutate(New_Country = na_if(Country, "Unspecified"), .keep = "used")
```

Its functioning relies on matchmaking as, when two vectors share the same values at the same positions, those values become NAs.

```{r, df_print = "paged"}
na_if(c(1, 2, 3, 4, 5), 
      c(5, 4, 3, 2, 1))
na_if(c(1, 2, 2, 3), 
      c(3, 2, 2, 1))
```

The documentation states that the second vector is changed to the type of the first, but I was only able to make it work with the easy conversion between TRUE/FALSE and 1/0.

```{r, error = TRUE, df_print = "paged"}
na_if(0:1, c(FALSE, TRUE))
na_if(c(FALSE, TRUE), 0:1)
na_if(as.character(1:5), 1:5)
```

The example at the start works because the value "Unspecified" is recycled to the size of the filtered `df`, but we can recycle only vectors of length 1, so if, in a data frame, we want to substitute several values the size of the column and the length of the vector used in `na_if()` must match.

```{r}
df %>%
  slice(71:72) %>%
  mutate(New_Country = na_if(Country, c("United Kingdom", "France")), .keep = "used")
```

`na_if()` works with columns' names, not positions.

```{r, error = TRUE}
df %>%
  mutate(Country = na_if(8, "Unspecified"))
```

We can use expressions inside it though.

```{r}
df %>%
  mutate(Price_Rank = na_if(min_rank(Price), 1)) %>%
  count(Price_Rank) %>%
  arrange(!is.na(Price_Rank))
```

# - *with group_by()*

A grouped data frame doesn't change the functioning of `na_if()`.

```{r}
df %>%
  group_by(Country) %>%
  mutate(Country = na_if(Country, "Unspecified"), .keep = "used")
```