---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged_rowwise")
```

```{r, echo=FALSE}
is_rowwise_df <- function(x) {
  inherits(x, "rowwise_df")
}
info_paged_rowwise_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  rowwise_info <- paste0("<div class=\"group-info\">Rowwise: ", 
                      paste0(group_vars(x), collapse = ", "), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (is_rowwise_df(x)) { 
    tab_info_rowwise <- paste0("<div class=\"info\">", tibble_info, " ", rowwise_info, "</div>")
    cat(tab_info_rowwise)
  } else if (dplyr::is_grouped_df(x)){
    tab_info_grouped <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info_grouped)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))    
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged_rowwise") {
    options$render = info_paged_rowwise_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

`nest_by()` shares some similarities with `group_by()`, as both group multiple rows of a data frame by the unique values of the columns specified as arguments.

```{r}
df %>%
  group_by(Country)
```

The difference is in the output as `nest_by()` creates a list-column called data, where every element of it is a data frame containing the rows pertaining to a unique value of the nesting column

```{r}
df %>%
  nest_by(Country)
```

or to every existing combination of them in case we specify more than one.

```{r}
df %>%
  nest_by(Country, `Customer ID`)
```

Notice how the rows are ordered by the nesting columns.

It is therefore similar to `group_nest()`,

```{r}
df %>%
  group_nest(Country)
```

besides the fact that `nest_by()` returns a rowwise data frame, so we can subsequently apply the same manipulations to each data frame independently.

```{r}
df %>%
  nest_by(Country) %>%
  reframe(Values = dim(data)) %>%
  bind_cols(tibble(Dimension = rep(c("Rows", "Columns"), 40)))
```

If we want to go back to the original data, we can use `unnest()` from `tidyr`.

```{r}
df %>%
  nest_by(Country) %>%
  tidyr:::unnest(cols = c(data))
```

The order of the columns has changed though (the nesting column has been moved to the front), the rows' order is the one of the nesting column plus we obtained a grouped data frame.

In case we want to change the nesting column, we have to apply `unnest()`, `ungroup()` and then `nest_by()` again.

```{r}
df %>%
  nest_by(Country) %>%
  tidyr:::unnest(cols = c(data)) %>%
  ungroup() %>%
  nest_by(`Customer ID`)
```

As it is not possible to override it as we can with `group_by()`.

```{r, error = TRUE}
df %>%
  group_by(Country) %>%
  group_by(`Customer ID`)
df %>%
  nest_by(Country) %>%
  nest_by(`Customer ID`)
```

Unless it is by the same columns, but that is redundant.

```{r}
df %>%
  nest_by(Country) %>%
  nest_by(Country)
df %>%
  nest_by(Country, `Customer ID`) %>%
  nest_by(Country, `Customer ID`)
```

To unnest we can also use `reframe()`, which ungroups the output as well.

```{r}
df %>%
  nest_by(Country) %>%
  reframe(data)
```

We can access every data frame like this,

```{r}
df %>%
  nest_by(Country) %>%
  ungroup() %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

also without selecting data if we want to show the nesting column as well.

```{r}
df %>%
  nest_by(Country) %>%
  ungroup() %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

Without ungrouping we would not get the desired output but we would unnest the whole data frame.

```{r}
df %>%
  nest_by(Country) %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

Like `group_by()`, we can use `nest_by()` with expressions (the documentation states that "Computations are not allowed in `nest_by()`" though).

```{r}
df %>%
  nest_by(CustomerID = as.character(`Customer ID`))
df %>%
  nest_by(Price_Rank = dense_rank(Price))
```

Which are equivalent to performing a `mutate()` call beforehand.

```{r}
df %>%
  mutate(CustomerID = as.character(`Customer ID`)) %>%
  nest_by(CustomerID)
df %>%
  mutate(Price_Rank = min_rank(Price)) %>%
  nest_by(Price_Rank)
```

# - *.key*

We can change the name of the list-column with the `.key` argument

```{r}
df %>%
  nest_by(Country, .key = "list of dfs")
```

# - *.keep*

With the `.keep` argument instead we control whether or not to keep the nesting column in the data frames.

```{r}
df %>%
  nest_by(Country , .keep = TRUE) %>%
  ungroup() %>%
  select(data) %>%
  slice(1) %>%
  tidyr::unnest(cols = c(data))
```

Ultimately, `nest_by()` is equivalent to these lines of code.

```{r}
df %>%
  group_by(Country) %>%
  summarise(data = list(pick(everything()))) %>%
  rowwise()
```

# - *with group_by()*

On a grouped data frame, `nest_by()` only works without an argument, as it inherits the grouping column from `group_by()`.

```{r, error = TRUE}
df %>%
  group_by(Country) %>%
  nest_by()
df %>%
  group_by(Country) %>%
  nest_by(`Customer ID`)
```

Also if we specify the same column.

```{r, error = TRUE}
df %>%
  group_by(Country) %>%
  nest_by(Country)
```