---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *fundamentals*

tidy-select

`pull()` is used to extract one single column from our data frame

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  pull(Country)
```

in the form of a vector.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  pull(Country) %>%
  class()
```

Beside the name, we can also use the position.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  pull(-1)
```

We can use it with some expressions. 

```{r, error = TRUE, df_print = "paged"}
df %>%
  slice(1:100) %>%
  pull(dense_rank(Price))
df %>%
  slice(1:100) %>%
  pull(CustomerID = `Customer ID`)
df %>%
  slice(1:100) %>%
  pull(as.Date(InvoiceDate, origin = "1970-01-01"))
```

But not selection statements though.

```{r, error = TRUE, df_print = "paged"}
df %>%
  mutate(Country_Factor = as.factor(Country)) %>%
  slice(1:100) %>%
  pull(pick(where(is.factor)))
```

Without an argument, it uses the last column, based on the idea that it was the most recently subject to manipulations (`mutate()` adds columns at the last position by default for example), hence the one we are currently interested in.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  pull()
```

# - *name*

We can specify another column to name the vector we extract with the optional argument `name`.

```{r, df_print = "paged"}
df %>% 
  slice(1:100) %>%
  pull(Country, name = `Customer ID`)
```

But that can't be an external one.

```{r, error = TRUE, df_print = "paged"}
cusnames <- df$`Customer ID`
df %>% 
  pull(Country, name = cusnames)
```

The `name` argument can be not named, but attention that that doesn't mean that we can extract two columns.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  pull(Country, `Customer ID`)
```

# - *with group_by()*

Grouping columns are ignored.

```{r, df_print = "paged"}
df %>%
  slice(1:100) %>%
  group_by(`Customer ID`) %>%
  pull(Country)
```