---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

# - *select()*

tidy-select

`select()` preserves the columns specified as its arguments, removing the other ones.

```{r}
df %>%
  select(Invoice)
df %>%
  select(Invoice, InvoiceDate)
```

If we specify a column that doesn't exist, the function will throw an error.

```{r, error = TRUE}
df %>%
  select(Invoice, Shipment)
```

The columns can be specified either by their names (quoted or unquoted)

```{r}
df %>%
  select("Invoice", InvoiceDate, `Customer ID`)
```

or by their positions (which means their column number).

```{r}
df %>%
  select(1, 5, 7)
```

The order by which we specify them will reflect on the output.

```{r}
df %>%
  select(InvoiceDate, Invoice, `Customer ID`)
```

The minus (`-`) and NOT (`!`) operators remove the columns specified.

```{r}
df %>%
  select(-Description)
df %>%
  select(!Description)
```

NOT doesn't work if applied to more than one column. 

```{r}
df %>%
  select(-Invoice, -StockCode)
df %>%
  select(!Invoice, !StockCode)
```

Unless we use it on a vector.

```{r}
df %>%
  select(!c(Invoice, StockCode))
```

By using a colon (`:`) we can preserve and remove consecutive columns.

```{r}
df %>%
  select(Invoice:Description)
df %>%
  select(1:3)
df %>%
  select(-(Invoice:Description))
df %>%
  select(!(Invoice:Description))
df %>%
  select(-(1:3))
df %>%
  select(!(1:3))
```

There are many helpers we can take advantage of;

## - *string matching helpers*

like the ones that do string matching on the columns' names,

```{r}
df %>%
  select(starts_with("I"))
df %>%
  select(ends_with("E"))
df %>%
  select(contains("voice"))
```

all of them not case sensitive by default (but that can be changed by setting the optional argument `ignore.case` to FALSE).

```{r}
df %>%
  select(contains("Voice", ignore.case = FALSE))
```

Another optional argument is `vars`, by which we can supply different columns to select from than the ones in the data frame.

```{r}
df %>%
  select(contains("Voice", vars = c("Invoice", "StockCode")))
```

We also have `matches()`, that employs regex logic instead, and has a further optional argument, `perl`, to use Perl-compatible regexps.

```{r}
df %>%
  select(matches("Voice"))
```

If our condition doesn't return any column, the output will be a data frame with no columns.

```{r}
df %>%
  select(starts_with("Z"))
```

## - *any_of()` & `all_of()*

`any_of()` and `all_of()` must be used when we have columns' names stored in an external vector.

```{r}
cols_vec <- c("Invoice", "InvoiceDate")
df %>%
  select(any_of(cols_vec))
df %>%
  select(all_of(cols_vec))
```

As without them `select()` will throw a warning.

```{r}
df %>%
  select(cols_vec)
```

Their difference comes off when a column from the vector is not present in the data frame, as `all_of()`throws an error.

```{r, error = TRUE}
cols_vec <- c("Invoice", "InvoiceDate", "Shipment")
df %>%
  select(all_of(cols_vec))
```

Where `any_of()` returns just the columns of the data frame that match the ones in the vector.

```{r}
df %>%
  select(any_of(cols_vec))
```

The same behavior is present in case of de-selections.

```{r, error = TRUE}
df %>%
  select(-all_of(cols_vec))
df %>%
  select(-any_of(cols_vec))
```

## - *num_range()*

`num_range()` can be useful in case we have columns that differ only for a number, like for instance `Description1`, `Description2`, ... , to produce more concise code.

```{r}
df %>%
  select(num_range("Description", 3:5))
```

As before, when there are no columns matching the condition, `select()` returns 0 rows instead of an error.

`num_range()` has the optional arguments `width`, with which we can account for a serie of 0s before the actual number (like if the columns were named `Description001`, `Description002`, ... , we would use a `width` of 3) and `vars`.   
Furthermore, we can specify `suffix =` if the columns were named `1Description`, `2Description`, ... ,

## - *last_col()*

`last_col()` lets us easily select the last

```{r}
df %>%
  select(last_col())
```

or the nth column from the end.

```{r}
df %>%
  select(last_col(5))
```

## - *where()*

Using `where()`, `select()` preserves the columns that evaluate to TRUE to the condition expressed within.

```{r}
df %>%
  select(where(is.numeric))
df %>%
  select(where(is.character))
```

We can rewrite those expressions in different ways; with an anonymous function,

```{r}
df %>% 
  select(where(function(x) is.numeric(x)))
```

or with a formula.

```{r}
df %>% 
  select(where(~ is.numeric(.x)))
```

Using a formula we can easily add multiple conditions,

```{r}
df %>%
  select(where(~ is.numeric(.x) & median(.x, na.rm = TRUE) < 50))
```

or more particular expressions, like selecting the columns that have at least one 12.

```{r}
df %>%
  select(where(~ any(.x == 12, na.rm = TRUE)))
```

Or where all the values are either "Italy" or "France".

```{r}
df %>%
  mutate(Country = Country,
         IT_Country = rep("Italy"),
         FR_Country = rep("France"),
         UK_Country = rep("United Kingdom"), .keep = "used")
df %>%
  mutate(Country = Country,
         IT_Country = rep("Italy"),
         FR_Country = rep("France"),
         UK_Country = rep("United Kingdom"), .keep = "used") %>%
  select(where(~ all(.x %in% c("Italy", "France"))))
```

We can use this syntax to preserve/remove columns with/without  NAs.

```{r}
df %>%
  select(where(~ any(is.na(.x))))
df %>%
  select(where(~ !any(is.na(.x))))
```

In case there were columns with only NAs,

```{r}
df %>%
  mutate(NA_column_1 = NA,
         NA_column_2 = NA, .before = "Invoice")
```

We can use `all()` to either preserve or remove them.

```{r}
df %>%
  mutate(NA_column = NA,
         NA_column_2 = NA) %>%
  select(where(~ all(is.na(.x))))
df %>%
  mutate(NA_column = NA,
         NA_column_2 = NA) %>%
  select(where(~ !all(is.na(.x))))
```

## - *with & (AND) & | (OR)*

We can chain more selections using the logical operators AND (`&`) and OR (`|`).

`&` returns the columns present in all selections' statements.

```{r}
df %>%
  select(c("Invoice", "Country") & last_col())
```

While `|` the ones present in any one of them.

```{r}
df %>%
  select(c("Invoice", "Country") | last_col())
```

Attention that, differently from `filter()`, here the comma is not synonymous with `&` but with `|`.

```{r}
df %>%
  select(c("Invoice", "Country"), last_col())
```

So for clarity's sake it is better to favor `|` to it.

```{r}
df %>%
  select(contains(c("A", "d")))
df %>%
  select(contains("A") | contains("d"))
df %>%
  select(contains("A") & contains("d"))
```

# - *rename()*

tidy-select

We can use `select()` also to rename columns, with the syntax `new_name = old_name`.

```{r}
df %>%
  select(CustomerID = `Customer ID`)
```

It will generally drop all unspecified columns though, except in the very specific case where we remove the old column before giving its name to another one.

```{r}
df %>%
  select(-Invoice, Invoice = StockCode)
df %>%
  select(!Invoice, Invoice = StockCode)
```

So it is better to use `rename()`, which preserves the unspecified ones.

```{r}
df %>%
  rename(CustomerID = `Customer ID`)
```

## - *any_of() & all_of()*

`any_of()` and `all_of()` can be used as well, when we have a vector with the modifications.

```{r}
any_new_names <- c(CUSTOMER_ID = "Customer ID", COUNTRY = "Country", SHIPMENT = "Shipment")
df %>%
  rename(any_of(any_new_names))
all_new_names <- c(CUSTOMER_ID = "Customer ID", COUNTRY = "Country")
df %>%
  rename(all_of(all_new_names))
```

By the same token as with `select()`, all of the columns to be modified must be present in the data frame when we use `all_of()`.

```{r, error = TRUE}
df %>%
  rename(all_of(any_new_names))
```

## - *rename_with()*

In case we want to use a function, one that changes the case for example, we can with its variant `rename_with()`, which first requires the function and then the columns we want to apply that function to.

```{r}
df %>%
  rename_with(stringr::str_to_lower, starts_with("I"))
```

The default is `everything()`, i.e. all of the columns.

```{r}
df %>%
  rename_with(stringr::str_to_lower)
```

## - *everything()*

`everything()` is in fact a helper that selects all of the columns, allowing to type less.

It can be useful if we want to move some columns to the first positions in a concise way, as it permits to not having to write the columns' name that we do not want to move.

```{r}
df %>%
  select(`Customer ID`, Country, everything())
```

It doesn't work though if we want to move columns at the end

```{r}
df %>%
  select(everything(), Invoice, `Customer ID`)
```

or if we use it between two columns. 

```{r}
df %>%
  select(Invoice, everything(), `Customer ID`)
```

Plus I don't understand the functioning of its `vars` argument.

```{r}
df %>%
  select(everything(vars = c("Description", "InvoiceDate")))
df %>%
  select(Invoice, everything(vars = c("Description", "InvoiceDate")), `Customer ID`)
```

# - *relocate()*

tidy-select

So, to move groups of columns, `relocate()` comes handy. 

```{r}
df %>%
  relocate(`Customer ID`, Country)
```

The default behavior is to move the columns specified at the front, but we can use its `.after` and `.before` optional tidy-select arguments to change that.

```{r}
df %>%
  relocate(Invoice, `Customer ID`, .after = Country)
```

`relocate()` employs many behaviors of `select()` and `rename()`.

```{r}
df %>% 
  relocate(where(is.numeric), .after = last_col())
df %>% 
  relocate(any_of(cols_vec))
df %>%
  relocate(CustomerID = `Customer ID`)
```

The optional arguments can utilize `where()` as well.

```{r}
df %>% 
  relocate(where(is.numeric), .after = where(is.character))
```

# - *with group_by()*

In case of grouped data frames the grouping columns are always added to the selection.

```{r}
df %>%
  group_by(Country) %>%
  select(Invoice, Quantity)
```

And can't be de-selected even manually.

```{r}
df %>%
  group_by(Country) %>%
  select(Invoice, Quantity, -Country)
```

The order of the selection of the previous example is important because, if we put `-Country` first, it will maintain all the columns.

```{r}
df %>%
  group_by(Country) %>%
  select(-Country, Invoice, Quantity)
```

Like if we put it last with a NOT (`!`).

```{r}
df %>%
  group_by(Country) %>%
  select(Invoice, Quantity, !Country)
```

The only way to remove the grouping columns from the selection is to `ungroup()`.

```{r}
df %>%
  group_by(Country) %>%
  ungroup() %>%
  select(Invoice, Quantity)
```

## - *group_cols()*

We can use `group_cols()` if we want to change the position of the grouping columns in the output, especially when we have several of them.

```{r}
df %>%
  group_by(Country, `Customer ID`) %>%
  select(Invoice, Quantity)
df %>%
  group_by(Country, `Customer ID`) %>%
  select(Invoice, Quantity, group_cols())
df %>%
  group_by(Country, `Customer ID`) %>%
  select(Invoice, group_cols(), Quantity)
```

Using it will also remove the message.

```{r}
df %>%
  group_by(Country) %>%
  select(group_cols())
df %>%
  group_by(Country) %>%
  select()
```

It doesn't work though for removing the grouping columns from the selection.

```{r}
df %>%
  group_by(Country) %>%
  select(Invoice, Quantity, -group_cols())
```

And if put first it will invalidate the `select()` call, like we have seen before.

```{r}
df %>%
  group_by(Country) %>%
  select(-group_cols(), Invoice, Quantity)
```

`rename()` and `relocate()` don't change their behavior with a grouped data frame,

```{r}
df %>%
  group_by(Country) %>%
  rename(CustomerID = `Customer ID`)
df %>%
  group_by(Country) %>%
  relocate(`Customer ID`, Country)
```

`rename()` though updates the name of the grouping columns when used on them.

```{r}
df %>%
  group_by(Country) %>%
  rename(COUNTRY = Country)
```