---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr)
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
```

data-masking

# - *fundamentals*

with `slice()` we can specify one or several row indexes and the function will preserve the relative rows.

```{r}
df %>%
  slice(5)
df %>%
  slice(5, 6)
```

If we want to preserve an interval of rows we can use a colon (`:`).

```{r}
df %>%
  slice(5:10)
```

If the interval specified exceeds the total number of rows, `slice()` will not throw an error.

```{r}
df %>%
  slice(500000:1000000)
```

`slice()` is equivalent to the following expressions (which must be used in case of relational databases).

```{r}
df %>%
  filter(row_number() == 5)
df %>%
  filter(row_number() %in% c(5, 6))
df %>%
  filter(row_number() %in% 5:10)
df %>%
  filter(row_number() %in% 500000:1000000)
```

`slice()` is very versatile and can perform several additional actions:

- remove one or several rows with the use of a minus (`-`)

```{r}
df %>%
  slice(-5)
df %>%
  slice(-5, -6)
df %>%
  slice(-(5:10))
```

- change the order of the output

```{r}
df %>%
  slice(10:5)
```

or duplicate rows,

```{r}
df %>%
  slice(7, 7, 7, 7)
```

even with complex expressions,

```{r}
df %>%
  slice(rep(c(1, 2, 3), c(1, 2, 4)))
```

as long as all the input indexes have all the same sign though.

```{r, error = TRUE}
df %>%
  slice(7, 7, 7, -7)
```

- by using `n()` (or `nrow(df)`, it is indifferent here), we can preserve or remove the last row or a number of rows from the bottom.

```{r}
df %>%
  slice(n())
df %>%
  slice(-n())
df %>%
  slice(525000:n())
df %>%
  slice(-(525000:n()))
df %>% 
  slice(1:(n() - 462))
```

As we've seen in all of these examples, if a row index is present in the expression, that row index is preserved (or removed when we used a `-`).

```{r}
df %>%
  slice(5:10)
```

We can think of it as if `slice()` uses >= / <= like `between()`.

```{r}
df %>%
  filter(between(row_number(), 5, 10))
```

Furthermore, as `slice()` is a data-masking function, it is possible to provide the indexes with calculations or with the output of a function, like we did in some of the previous examples.

```{r}
df %>%
  slice(c(2, 4, 6, 8) / 2)
df %>%
  slice(seq(1, 8, by = 2))
```

With no argument, `slice()` returns an empty data frame.

```{r}
df %>%
  slice()
```

`slice()` has many helpers that can streamline the code in particular cases:

# - *slice_head() & slice_tail()*

`slice_head()` and `slice_tail()` respectively preserve the first and last `n` rows. `n` must be an integer.

```{r}
df %>%
  slice_head(n = 5)
df %>%
  slice_tail(n = 5)
```

Keep in mind that `n` must always be explicit with the helpers.

```{r, error = TRUE}
df %>%
  slice_head(5)
```

While `slice()` would throw an error if we used it.

```{r, error = TRUE}
df %>%
  slice(n = 5)
```

The default is `n` = 1.

```{r}
df %>%
  slice_head()
```

The example with `slice_tail()` can be rewritten like this, but it is surely more convoluted, so the presence of this helper is appreciated.

```{r}
n <- 5
df %>%
  slice((n() - (n - 1)):n())
```

A similar expression, not possible with `slice_tail()`, could be used though to select a range of row indexes `n` to `n` + `n` counting from the bottom.

```{r}
df %>% 
  slice(((n() - (n - 1)) - n):(n() - (n - 1))) 
```

With `slice_head()` and `slice_tail()` we can also use the `prop` argument (that needs to be explicit, like `n`) to preserve a user defined proportion of rows.

```{r}
df %>%
  slice_head(prop = 0.1)
df %>%
  slice_tail(prop = 0.1)
```

The number of rows is rounded down in case of decimals in the output.

```{r}
df %>%
  slice_head(prop = 0.6)
```


We can use a minus (`-`), like we do in `slice()`, but the result is counter intuitive in my opinion as `slice_head()` removes the rows at the bottom and `slice_tail()` the ones at the top.

```{r}
df %>%
  slice_head(n = -5)
df %>%
  slice_tail(n = -5)
df %>%
  slice_head(prop = -0.1)
df %>%
  slice_tail(prop = -0.1)
```

# - *slice_sample()*

`slice_sample()` samples a number or a percentage of rows.
Like with the previous helpers, `n` and `prop` must always be explicit, `n` an integer and the number of rows is rounded down if the output contains decimals when using `prop`.

```{r}
df %>%
  slice_sample(n = 5)
df %>%
  slice_sample(prop = 0.1)
```

The default is again `n` = 1.

```{r}
df %>%
  slice_sample()
```

In case we want to remove random rows, we can provide negative integers to `n` and `prop`.

```{r}
df %>%
  slice_sample(n = -5)
df %>%
  slice_sample(prop = -0.1)
```

## - *replace*

The `replace` argument controls whether the sampling occurs without (the default FALSE) or with replacements (which means that a row can be picked more than once).

```{r}
df %>%
  slice(1:5)
df %>%
  slice(1:5) %>%
  slice_sample(n = 5, replace = TRUE)
```

## - *weight_by*

data-masking

If we want to give more representation to a certain segment of the data, we can use the `weight_by` argument, which increases the probability of preserving rows with higher values in the column specified in it.

In this example we weight by the `Price` column, so expensive items will be more present in the output compared to a non weighted sample. We can check it with the `Price_Rank` column, whose lower values indicates a more expensive item.

```{r}
df %>%
  filter(Price >= 0) %>%
  mutate(Price_Rank = dense_rank(desc(Price))) %>%
  slice_sample(n = 10, weight_by = Price)
df %>%
  filter(Price >= 0) %>%
  mutate(Price_Rank = dense_rank(desc(Price))) %>%
  slice_sample(n = 10)
```

Notice the `filter()` call on the column we provide to `weight_by`, as that must be non-negative. We can also use an external vector, as long as it is of the same size as the number of rows of the data frame fed into `slice_sample()`.

We can't in fact use inline subsetting on `weight_by` because it will produce a different one.

```{r, error = TRUE}
df %>%
  mutate(Price_Rank = dense_rank(desc(Price))) %>%
  slice_sample(n = 10, weight_by = Price[Price > 0])
```

We can use expressions inside `weight_by`.

```{r}
df %>%
  filter(Price >= 0) %>%
  mutate(Price_Rank = dense_rank(desc(Price))) %>%
  slice_sample(n = 10, weight_by = Price / 2)
```

# - *slice_max() & slice_min()*

Last helpers, `slice_max()` and `slice_min()` preserve a number of rows with the maximum or minimum values of the column specified in the `order_by` argument, sorting the output as well.

```{r}
df %>%
  slice_max(order_by = Quantity, n = 10)
df %>%
  slice_min(order_by = Quantity, n = 10)
```

Being a data-masking argument, `order_by` accepts expressions,

```{r}
df %>%
  slice_max(order_by = Quantity / 2, n = 10)
```

and can be implicit.

```{r}
df %>%
  slice_max(Quantity, n = 10)
```

But `n` cannot,

```{r, error = TRUE}
df %>%
  slice_max(Quantity, 10)
```

and, as with other helpers, its default is 1 and it accepts only integer values.

```{r}
df %>%
  slice_max(Quantity)
```

In the presence of ties all of the rows are preserved and we assume, as with `arrange()`, that the rows are ordered by their original row indexes.

```{r}
df %>%
  slice_min(Quantity, n = 4)
```

To break the ties we can either specify two columns, that must be in a data frame or tibble format, so for example we can use `pick()` to do that,

```{r}
df %>%
  slice_min(pick(Quantity, StockCode), n = 4)
```

## - *with_ties*

or we can set the `with_ties` argument to FALSE, so the function will remove the rows with an output index superior to `n`.

```{r}
df %>%
  slice_min(Quantity, n = 4, with_ties = FALSE)
```

We remind that, within ties, the rows of the output are ordered by their original row index so if we are not satisfied by which row gets removed we can use `arrange()` before slicing, to preserve invoice 530348 instead.

```{r}
df %>%
  arrange(desc(InvoiceDate)) %>%
  slice_min(Quantity, n = 4, with_ties = FALSE)
```

NAs are included if they are necessary to reach the number specified in `n` or `prop`.

```{r}
df %>%
  filter(StockCode == "90081B")
df %>%
  filter(StockCode == "90081B") %>%
  slice_max(`Customer ID`, n = 3)
```

## - *na_rm*

With `na_rm = TRUE` we can exclude them, it might mean returning less rows than specified in `n` though.

```{r}
df %>%
  filter(StockCode == "90081B") %>%
  slice_max(`Customer ID`, n = 3, na_rm = TRUE)
```

We can obtain the same results provided by `slice_min()` or `slice_max()` with a combination of `arrange()` plus `slice()`.

```{r}
df %>%
  slice_min(Quantity, n = 10)
df %>%
  arrange(Quantity) %>%
  slice(1:10)
```

Or by using `filter()` with a ranking function.
The order will be different here as `filter()` preserves the original row index's one.

```{r}
df %>%
  filter(row_number(Quantity) <= 10)
```

If we wrap the variable with `desc()`, we can invert the behavior.

```{r}
df %>%
  slice_max(desc(Quantity), n = 10)
df %>%
  slice_min(desc(Quantity), n = 10)
```

As with the other helpers, a proportion of rows can be used as well and they are rounded down in case of decimals.

```{r}
df %>%
  slice_max(Quantity, prop = 0.1)
df %>%
  slice_min(Quantity, prop = 0.1)
```

Just like `n`, `prop` must be explicit.

```{r, error = TRUE}
df %>%
  slice_min(Quantity, 0.1)
```

`slice_max()` and `slice_min()` can also be used with other classes of columns:

- with character ones we will rank by the alphabet, as it is defined in your locale.

```{r}
df %>%
  slice_min(Description, n = 10)
```

- with factors by the order of the levels.

```{r}
df %>%
  mutate(Description_Factor = factor(Description, levels = sample(unique(df$Description), 
                                                                  length(unique(df$Description))), ordered = TRUE)) %>%
  slice_min(Description_Factor, n = 1000)
```

- with datetime objects it depends on the class, with for example POSIXct will be by the number of seconds since 01/01/1970 (in the UTC time zone) as a numeric vector. This is quite resource heavy so I computed on only the first 100 rows.

```{r}
df %>%
  slice(1:100) %>%
  slice_min(InvoiceDate, n = 10)
```

A minus before `n` doesn't seem to affect the output.

```{r}
df %>%
  slice(21:30)
df %>%
  slice(21:30) %>%
  slice_min(Quantity, n = 5)
df %>%
  slice(21:30) %>%  
  slice_min(Quantity, n = -5)
```

# - *with group_by() / .by*

`slice()` and its helpers work as expected with grouped data frames, preserving as many specified rows indexes, `n` or `prop` as there are groups. 
It is important to notice that they rearrange the output by the grouping column, like `summarise()` and unlike `filter()`.

```{r}
df %>%
  group_by(Country) %>%
  slice(2)
df %>%
  group_by(Country) %>%
  slice_head(n = 5)
```

As with the ungrouped case, when we use a proportion the number of rows we preserve is rounded down.

```{r}
df %>%
  group_by(Country) %>%
  slice(1:5)
df %>%
  group_by(Country) %>%
  slice(1:5) %>%
  slice_head(prop = 0.5)
```

And as already seen at the beginning, `slice()` doesn't throw an error if the interval specified crosses over the number of rows.

```{r}
df %>%
  group_by(Country) %>%
  slice(11:15)
```

## - *.preserve*

`slice()` (but not its helpers) has a `.preserve` argument that, when set to TRUE, doesn't recalculate the number of groups when the output doesn't have any rows from a particular one (Lebanon here).

```{r, df_print = "paged"}
df %>%
  group_by(Country) %>%
  slice(15) %>%
  n_groups()
df %>%
  group_by(Country) %>%
  slice(15, .preserve = TRUE) %>%
  n_groups()
```

This is useful if further operations need the original number of groups (notice how Lebanon is present in the output even if it has no rows).

```{r}
df %>%
  group_by(Country) %>%
  slice(15, .preserve = TRUE) %>%
  summarise(n = n())
```

Besided `group_by()`, we can also use the `.by` argument to use `slice()` on grouped data frames.

```{r}
df %>%
  group_by(Country) %>%
  slice(2)
df %>%
  slice(2, .by = Country)
```

You can see that using `.by` the output maintains the original row order plus the data frame is returned ungrouped.

Another difference is that `.by` uses a tidy-select syntax, to keep in mind if we want to select more than one column.

```{r}
df %>%
  slice(2, .by = c("Country", "Customer ID"))
```

That is an useful feature to employ selection helpers.

```{r}
df %>%
  slice(2, .by = starts_with("C"))
```

The helpers can use this argument as well, but with them it is called `by`.

```{r}
df %>%
  slice_head(n = 5, by = Country)
```