---
title: ""
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, df_print="info_paged")
```

```{r, echo=FALSE}
info_paged_print <- function(x, options) {
  tibble_info <- paste0("<div class=\"tibble-info\">A tibble: ", nrow(x), " x ", ncol(x), "</div>")
  group_info <- paste0("<div class=\"group-info\">Groups: ", 
                      paste0(group_vars(x), collapse = ", "), 
                      " [", nrow(group_keys(x)), "]", "</div>")
  
  if (dplyr::is_grouped_df(x)) {
    tab_info <- paste0("<div class=\"info\">", tibble_info, " ", group_info, "</div>")
    cat(tab_info)
  } else {
    cat(paste0("<div class=\"info\">", tibble_info, "</div>"))
  }
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(x, "options")),
    meta = list(
      dependencies = rmarkdown:::html_dependency_pagedtable()
    )
  )
}

knitr::opts_hooks$set(df_print = function(options) {
  if (options$df_print == "info_paged") {
    options$render = info_paged_print
    options$comment = ""
    options$results = "asis"
  }
  options
})
```

```{css, echo=FALSE}
.tibble-info,
.group-info {
  display: inline-block;
  padding: 15px;
}

.info {
  margin-top: 5px;
  margin-bottom: 5px;
  border: 1px solid #ccc;
  border-radius: 4px;
  font-weight: 600;
  color: #999898;
}
```

```{r, message = FALSE, echo=FALSE}
library(dplyr) #OPPURE USO include?
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx")) 
```

# - *fundamentals*

`tally()` is a wrapper for `summarise(n = n())` and as such it returns the number of rows of a data frame.

```{r}
df %>%
  tally()
df %>% 
  summarise(n = n())
```

`tally()`, like `n()`, doesn't need an argument and if supplied with an unnamed one it uses it as the value for `wt`, so attention is required.

```{r}
df %>%
  tally(Quantity)
df %>%
  tally(wt = Quantity)
```

# - *wt*

data-masking

Speaking about `wt`, it modifies the function inside `summarise()` from `n()` to `sum(wt)`,

```{r}
df %>%
  tally(wt = Quantity)
df %>% 
  summarise(n = sum(Quantity))
```

Therefore outputting not the number of rows but the summation of the values of the column specified in it.

We can't obviously use `wt` with columns whose values can't be summed.

```{r, error = TRUE}
df %>%
  tally(wt = Description)
```

It can be used though with more than one column.

```{r}
df %>%
  tally(wt = c(Quantity, Price))
```

In these instances it is equivalent to

```{r}
df %>% 
  summarise(n = sum(Quantity, Price))
```

`wt` accepts expressions as well, 

```{r}
df %>%
  tally(wt = Quantity / 2)
```

as long as their output can be summed of course.

```{r, error = TRUE}
df %>%
  tally(wt = as.character(Quantity / 2))
```

There are some functions where the summation is inconsequential,

```{r}
df %>%
  tally(wt = n_distinct(`Customer ID`))
```

as they return just one addend.

```{r}
df %>%
  summarise(n = sum(n_distinct(`Customer ID`)))
df %>%
  summarise(n = n_distinct(`Customer ID`))
```

so using `wt` gives us the possibility to employ them in a `tally(`) call, if we wish to do so.

If we have an already aggregated data frame,

```{r}
df %>%
  count(Country)
```

we can use `wt` to retrieve the original total number of rows.

```{r}
df %>%
  count(Country) %>%
  tally(wt = n)
```

# - *name*

Another optional argument is `name`, to change `n` to a custom denomination.

```{r}
df %>%
  tally(name = "Total_Number_of_Rows")
```

# - *with group_by()*

When used on a grouped data frame, `tally()` returns the number of rows for each group thus becoming equivalent to a `count()` call,

```{r}
df %>%
  group_by(Country) %>%
  tally()
df %>%
  count(Country)
```

except from the fact that `tally()`, like `summarise()`, removes the most recent grouping when we have a data frame grouped by more than one column,

```{r, df_print = "paged"}
df %>%
  group_by(Country, `Customer ID`) %>%
  tally() %>%
  group_vars()
df %>%
  group_by(Country, `Customer ID`) %>%
  summarise(n = n()) %>%
  group_vars()
```

while `count()` removes all the groups.

```{r, df_print = "paged"}
df %>%
  count(Country, `Customer ID`) %>%
  group_vars()
```

## - *sort*

With a grouped data frame, we can use the `sort` argument to arrange the rows by `n`, in descending order.

```{r}
df %>%
  group_by(Country) %>%
  tally(sort = TRUE)
```

Using `sort  = TRUE` we can see as well that NAs are counted as one value, as per `n()` functionality.

```{r}
df %>%
  group_by(`Customer ID`) %>%
  tally(sort = TRUE)
```

# - *add_tally()*

`tally()`'s variant, `add_tally()`, uses `mutate()` instead of `summarise()` therefore adding a column named `n` with the same value for all of the rows.

```{r}
df %>%
  add_tally()
df %>%
  mutate(n = n())
```

`add_tally()` is consistent with the arguments of `tally()`.

```{r}
df %>%
  tally(wt = Quantity, name = "Total_Quantity")
```

## - *with group_by()*

On a grouped data frame all the rows pertaining to the same group will share the same value.

```{r}
df %>%
  group_by(Country) %>%
  add_tally()
```

Differently from `tally()`, `add_tally()` doesn't ungroup the output, even with several column, 

```{r}
df %>%
  group_by(Country, `Customer ID`) %>%
  add_tally()
```

as that is not a property of `mutate()`.

```{r}
df %>%
  group_by(Country, `Customer ID`) %>%
  mutate(n = n())
```

## - *sort*

With a grouped data frame we can use the `sort` argument, that will arrange the rows by the descending order of `n`.

```{r}
df %>%
  group_by(Country) %>%
  add_tally(sort = TRUE)
```